---
title: "14 Power analysis"
author: "Tobias Gerstenberg"
date: "2/8/2019"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

> In general, the class was good. However, there were several concepts that were not clearly explained. For example, what is the difference between p value and alpha? Can you write it in a probability formula to explain the differences? Also, for the contrasts in R, may be it is good to know the meaning of estimated marginal means? Are they different from calculating the mean for a particular group? Finally, part of the codes about contrasts are not in the lecture note.

> To counter my classmates, if there's no opportunity to ask questions then if you get lost, it's very hard to catch back up. The rest of the lecture would then become very hard to follow.

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-14.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

## What to do next time

- check how to do linear contrasts for unbalanced designs 
- explain what estimated marginal means are (to interpret estimates of `emmeans()`)
- try to clarify conceptual distinctions (e.g. $\alpha$ vs. p-value)
- make sure that none of the code is missing on the slides (remove the white boxes in the print version)
