---
title: "25 Model assumptions and reporting"
author: "Tobias Gerstenberg"
date: "3/9/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(theme_classic() + 
            theme(
              text = element_text(size = 20),
              panel.grid.major.y = element_line(color = "gray90")))
```

> Thank you for teaching this course. I really appreciate the environment you created within the course and within office hours; we felt challenged, but also felt that there was support for us. I also really appreciate the embedded openness and optimism that pervaded the course, as this can be a dry and challenging subject.

> Not related to this class in particular, but I would like to mention that, while I feel I understood frequentist statistics quite well, I don't feel the same way with Bayesian statistics. I would have liked to have more practice on Bayesian data analysis.

> Today's class was incredibly helpful and provided a comprehensive overview of everything we covered in the course. Thank you so, so much to the entire teaching team! :)

## Student feedback 

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-25.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read.csv("how-was-the-pace-of-25.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  scale_y_continuous(breaks = seq(0, 30, 2), labels = seq(0, 30, 2)) + 
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-25.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  scale_y_continuous(breaks = seq(0, 30, 2), labels = seq(0, 30, 2)) + 
  theme(title = element_text(size = 18))
```

## What to do next time

- try to improve question handling via zoom 
- maybe try and have one more class on Bayesian data analysis?! 


