---
title: "Approximating a Gaussian Distribution via Metropolis-Hastings"
author: "AI Assistant"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(ggplot2)
```

## Introduction

The Metropolis-Hastings (MH) algorithm is a Markov Chain Monte Carlo (MCMC) method for obtaining a sequence of random samples from a probability distribution from which direct sampling is difficult. This sequence can be used to approximate the distribution or to compute an integral (e.g., an expected value).

In this document, we illustrate the algorithm by using it to sample from a standard Gaussian (Normal) distribution, $\mathcal{N}(0, 1)$. While we can easily sample from a Gaussian directly using `rnorm()`, applying MH to this known distribution is a great way to understand how the algorithm works.

## Target Distribution

The target distribution is the standard normal distribution. Its unnormalized probability density function is proportional to:

$$ P(x) \propto \exp\left(-\frac{1}{2}x^2\right) $$

We define a function that returns this unnormalized density.

```{r target}
target_density <- function(x) {
  exp(-0.5 * x^2)
}
```

## The Metropolis-Hastings Algorithm

We will use a symmetric random walk proposal distribution, specifically a Normal distribution centered at the current state: $q(x' | x) = \mathcal{N}(x, \sigma^2)$.

Because the proposal distribution is symmetric (i.e., $q(x' | x) = q(x | x')$), the acceptance ratio simplifies to:

$$ \alpha = \min\left(1, \frac{P(x')}{P(x)}\right) $$

Here is the implementation of the algorithm:

```{r mh-algo}
metropolis_hastings <- function(n_iter, start_value, proposal_sd) {
  # Initialize vector to store samples
  samples <- numeric(n_iter)
  samples[1] <- start_value
  
  # Counter for accepted proposals
  accepted <- 0
  
  for (i in 2:n_iter) {
    current_x <- samples[i - 1]
    
    # 1. Propose a new state from the proposal distribution
    proposed_x <- rnorm(1, mean = current_x, sd = proposal_sd)
    
    # 2. Calculate the acceptance ratio
    # (Using the unnormalized target density)
    p_current <- target_density(current_x)
    p_proposed <- target_density(proposed_x)
    
    # We take the minimum of 1 and the ratio
    ratio <- p_proposed / p_current
    
    # 3. Accept or reject the proposal
    if (runif(1) < ratio) {
      samples[i] <- proposed_x
      accepted <- accepted + 1
    } else {
      samples[i] <- current_x
    }
  }
  
  message(sprintf("Acceptance rate: %.2f%%", (accepted / (n_iter - 1)) * 100))
  return(samples)
}
```

## Running the Algorithm

Let's run the algorithm for 10,000 iterations. We will start at an arbitrary value (e.g., `x = 5`) to observe how the chain converges to the target distribution, and we will use a proposal standard deviation of `1`.

```{r run-algo}
set.seed(42) # For reproducibility
n_iterations <- 10000
start_val <- 5.0
prop_sd <- 1.0

# Generate samples
samples <- metropolis_hastings(n_iter = n_iterations, 
                               start_value = start_val, 
                               proposal_sd = prop_sd)

# Create a data frame for plotting
df_samples <- data.frame(
  iteration = 1:n_iterations,
  value = samples
)
```

## Diagnostics and Results

### Trace Plot

A trace plot shows the sampled values over iterations. It helps us visualize the "mixing" of the chain and determine if it has converged to the stationary distribution. Notice how the chain quickly moves from the starting value of `5` towards the high-probability region around `0`.

```{r trace-plot}
ggplot(df_samples, aes(x = iteration, y = value)) +
  geom_line(color = "steelblue", alpha = 0.8) +
  labs(title = "Trace Plot of MCMC Samples",
       x = "Iteration",
       y = "Sample Value") +
  theme_minimal()
```

### Histogram vs. Target Density

To see if our samples successfully approximate the target distribution, we plot a histogram of the samples (discarding the first 1000 iterations as "burn-in") and overlay the true probability density function of the standard normal distribution.

```{r histogram}
# Discard the first 1000 samples as burn-in
burn_in <- 1000
samples_kept <- df_samples$value[(burn_in + 1):n_iterations]
df_kept <- data.frame(value = samples_kept)

ggplot(df_kept, aes(x = value)) +
  # Histogram of MCMC samples
  geom_histogram(aes(y = ..density..), bins = 50, fill = "lightblue", color = "white", alpha = 0.7) +
  # True Standard Normal Density
  stat_function(fun = dnorm, args = list(mean = 0, sd = 1), 
                color = "red", linewidth = 1.2, linetype = "dashed") +
  labs(title = "Approximated vs. True Gaussian Distribution",
       subtitle = paste("Red dashed line: True N(0,1) density\nHistogram: MH Samples (after", burn_in, "burn-in iterations)"),
       x = "Value",
       y = "Density") +
  theme_minimal()
```

As we can see, the histogram of the samples closely matches the theoretical density of the Gaussian distribution, demonstrating that the Metropolis-Hastings algorithm successfully approximated the target distribution.
