---
title: "Class 18"
author: "Tobias Gerstenberg"
date: "February 16th, 2022"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=17"]
---

# Linear mixed effects models 2

## Learning goals 

- An `lmer()` worked example
  - complete pooling vs. no pooling vs. partial pooling
  - getting p-values 
  - checking model assumptions
- Simulating mixed effects models
  - effect of outliers 
  - non-homogeneity of variance 
  - Simpson's paradox 

## Load packages and set plotting theme  

```{r, message=FALSE, warning=FALSE}
library("knitr")       # for knitting RMarkdown 
library("kableExtra")  # for making nice tables
library("janitor")     # for cleaning column names
library("broom.mixed") # for tidying up linear models 
library("ggeffects")   # for plotting marginal effects
library("emmeans")     # for the joint_tests() function 
library("lme4")        # for linear mixed effects models
library("performance") # for assessing model performance
library("see")         # for assessing model performance
library("tidyverse")   # for wrangling, plotting, etc. 
```

```{r, warning=FALSE}
theme_set(theme_classic() + #set the theme 
            theme(text = element_text(size = 20))) #set the default text size

opts_chunk$set(comment = "",
               fig.show = "hold")
```

## A worked example

Let's illustrate the concept of pooling and shrinkage via the sleep data set that comes with the lmer package. 

```{r}
# load sleepstudy data set 
df.sleep = sleepstudy %>% 
  as_tibble() %>% 
  clean_names() %>% 
  mutate(subject = as.character(subject)) %>% 
  select(subject, days, reaction)
```

```{r}
# add two fake participants (with missing data)
df.sleep = df.sleep %>% 
  bind_rows(tibble(subject = "374",
                   days = 0:1,
                   reaction = c(286, 288)),
            tibble(subject = "373",
                   days = 0,
                   reaction = 245))
```

Let's start by visualizing the data 

```{r}
# visualize the data
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

The plot shows the effect of the number of days of sleep deprivation on the average reaction time (presumably in an experiment). Note that for participant 373 and 374 we only have one and two data points respectively. 

### Complete pooling 

Let's first fit a model the simply combines all the data points. This model ignores the dependence structure in the data (i.e. the fact that we have repeated observations from the same participants). 

```{r}
fit.complete = lm(formula = reaction ~ days,
                  data = df.sleep)

fit.params = tidy(fit.complete)

summary(fit.complete)
```

And let's visualize the predictions of this model.

```{r}
# visualization (aggregate) 
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_abline(intercept = fit.params$estimate[1],
              slope = fit.params$estimate[2],
              color = "blue") +
  geom_point() +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

And here is what the model's predictions look like separated by participant.

```{r}
# visualization (separate participants) 
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_abline(intercept = fit.params$estimate[1],
              slope = fit.params$estimate[2],
              color = "blue") +
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

The model predicts the same relationship between sleep deprivation and reaction time for each participant (not surprising since we didn't even tell the model that this data is based on different participants). 

### No pooling 

We could also fit separate regressions for each participant. Let's do that.

```{r}
# fit regressions and extract parameter estimates 
df.no_pooling = df.sleep %>% 
  group_by(subject) %>% 
  nest(data = c(days, reaction)) %>% 
  mutate(fit = map(data, ~ lm(reaction ~ days, data = .)),
         params = map(fit, tidy)) %>% 
  ungroup() %>% 
  unnest(c(params)) %>% 
  select(subject, term, estimate) %>% 
  complete(subject, term, fill = list(estimate = 0)) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  clean_names()
```

And let's visualize what the predictions of these separate regressions would look like: 

```{r}
ggplot(data = df.sleep,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_abline(data = df.no_pooling %>% 
                filter(subject != 373),
              aes(intercept = intercept,
                  slope = days),
              color = "blue") +
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

When we fit separate regression, no information is shared between participants. 

### Partial pooling 

By usign linear mixed effects models, we are partially pooling information. That is, the estimates for one participant are influenced by the rest of the participants.

We'll fit a number of mixed effects models that differ in their random effects structure. 

#### Random intercept and random slope

This model allows for random differences in the intercepts and slopes between subjects (and also models the correlation between intercepts and slopes). 

Let's fit the model

```{r}
fit.random_intercept_slope = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                                  data = df.sleep)
```

and take a look at the model's predictions: 

```{r, warning=FALSE, message=F}
fit.random_intercept_slope %>% 
  augment() %>% 
  clean_names() %>% 
ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_line(aes(y = fitted),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

As we can see, the lines for each participant are different. We've allowed for the intercept as well as the relationship between sleep deprivation and reaction time to be different between participants. 

#### Only random intercepts 

Let's fit a model that only allows for the intercepts to vary between participants. 

```{r}
fit.random_intercept = lmer(formula = reaction ~ 1 + days + (1 | subject),
                            data = df.sleep)
```

And let's visualize what these predictions look like: 

```{r, warning=FALSE, message=F}
fit.random_intercept %>% 
  augment() %>% 
  clean_names() %>% 
ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_line(aes(y = fitted),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

Now, all the lines are parallel but the intercept differs between participants. 

#### Only random slopes 

Finally, let's compare a model that only allows for the slopes to differ but not the intercepts. 

```{r}
fit.random_slope = lmer(formula = reaction ~ 1 + days + (0 + days | subject),
                        data = df.sleep)
```

And let's visualize the model fit: 

```{r, warning=FALSE, message=F}
fit.random_slope %>% 
  augment() %>% 
  clean_names() %>% 
ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_line(aes(y = fitted),
            color = "blue") + 
  geom_point() +
  facet_wrap(vars(subject), ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

Here, all the lines have the same starting point (i.e. the same intercept) but the slopes are different. 

### Compare results 

Let's compare the results of the different methods -- complete pooling, no pooling, and partial pooling (with random intercepts and slopes). 

```{r, warning=F, message=F}
# complete pooling
fit.complete_pooling = lm(formula = reaction ~ days,
                          data = df.sleep)  

df.complete_pooling =  fit.complete_pooling %>% 
  augment() %>% 
  bind_rows(fit.complete_pooling %>% 
              augment(newdata = tibble(subject = c("373", "374"),
                                       days = rep(10, 2)))) %>% 
  clean_names() %>% 
  select(reaction, days, complete_pooling = fitted)

# no pooling
df.no_pooling = df.sleep %>% 
  group_by(subject) %>% 
  nest(data = c(days, reaction)) %>% 
  mutate(fit = map(data, ~ lm(reaction ~ days, data = .)),
         augment = map(fit, augment)) %>% 
  unnest(c(augment)) %>% 
  ungroup() %>% 
  clean_names() %>% 
  select(subject, reaction, days, no_pooling = fitted)

# partial pooling
fit.lmer = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                data = df.sleep) 

df.partial_pooling = fit.lmer %>% 
  augment() %>% 
  bind_rows(fit.lmer %>% 
              augment(newdata = tibble(subject = c("373", "374"),
                                       days = rep(10, 2)))) %>% 
  clean_names() %>% 
  select(subject, reaction, days, partial_pooling = fitted)

# combine results
df.pooling = df.partial_pooling %>% 
  left_join(df.complete_pooling,
            by = c("reaction", "days")) %>% 
  left_join(df.no_pooling,
            by = c("subject", "reaction", "days"))
```

Let's compare the predictions of the different models visually: 

```{r, warning=FALSE, message=F}
ggplot(data = df.pooling,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_smooth(method = "lm",
              se = F,
              color = "orange",
              fullrange = T) + 
  geom_line(aes(y = complete_pooling),
            color = "green") + 
  geom_line(aes(y = partial_pooling),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

To better see the differences between the approaches, let's focus on the predictions for the participants with incomplete data: 

```{r, warning=FALSE, message=F}
# subselection
ggplot(data = df.pooling %>% 
         filter(subject %in% c("373", "374")),
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_smooth(method = "lm",
              se = F,
              color = "orange",
              fullrange = T) + 
  geom_line(aes(y = complete_pooling),
            color = "green") + 
  geom_line(aes(y = partial_pooling),
            color = "blue") + 
  geom_point() +
  facet_wrap(vars(subject)) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

#### Coefficients 

One good way to get a sense for what the different models are doing is by taking a look at the coefficients: 

```{r}
coef(fit.complete_pooling)
```

```{r}
coef(fit.random_intercept)
```

```{r}
coef(fit.random_slope)
```

```{r}
coef(fit.random_intercept_slope)
```

#### Shrinkage 

In mixed effects models, the variance of parameter estimates across participants shrinks compared to a no pooling model (where we fit a different regression to each participant). Expressed differently, individual parameter estimates are borrowing strength from the overall data set in mixed effects models. 

```{r}
# get estimates from partial pooling model
df.partial_pooling = fit.random_intercept_slope %>% 
  coef() %>% 
  .$subject %>% 
  rownames_to_column("subject") %>% 
  clean_names()

# combine estimates from no pooling with partial pooling model 
df.plot = df.sleep %>% 
  group_by(subject) %>% 
  nest(data = c(days, reaction)) %>% 
  mutate(fit = map(data, ~ lm(reaction ~ days, data = .)),
         tidy = map(fit, tidy)) %>% 
  unnest(c(tidy)) %>% 
  select(subject, term, estimate) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  clean_names() %>% 
  mutate(method = "no pooling") %>% 
  bind_rows(df.partial_pooling %>% 
              mutate(method = "partial pooling")) %>% 
  pivot_longer(cols = -c(subject, method),
               names_to = "index", 
               values_to = "value") %>% 
  mutate(index = factor(index, levels = c("intercept", "days")))

# visualize the results  
ggplot(data = df.plot,
       mapping = aes(x = value,
                     group = method,
                     fill = method)) + 
  stat_density(position = "identity",
               geom = "area",
               color = "black",
               alpha = 0.3) +
  facet_grid(cols = vars(index),
             scales = "free")
```

### Getting p-values 

To get p-values for mixed effects models, I recommend using the `joint_tests()` function from the `emmeans` package.

```{r}
lmer(formula = reaction ~ 1 + days + (1 + days | subject),
     data = df.sleep) %>% 
  joint_tests()
```

Our good ol' model comparison approach produces a Likelihood ratio test in this case: 

```{r}
fit1 = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
            data = df.sleep)

fit2 = lmer(formula = reaction ~ 1 + (1 + days | subject),
            data = df.sleep)

anova(fit1, fit2)
```

### Reporting results

#### Plotting marginal effects 

```{r}
# library("ggeffects")

# using the plot() function
ggpredict(model = fit.random_intercept_slope,
          terms = "days",
          type = "fe") %>% 
  plot()

# using our own ggplot magic
df.plot = ggpredict(model = fit.random_intercept_slope,
                    terms = "days",
                    type = "fe")

ggplot(data = df.plot,
       mapping = aes(x = x, 
                     y = predicted,
                     ymin = conf.low,
                     ymax = conf.high)) + 
  geom_ribbon(fill = "lightblue") +
  geom_line(size = 1)
```

#### Checking model performance 

```{r, message=FALSE, warning=FALSE}
lmer(formula = reaction ~ 1 + days + (1 + days | subject),
     data = df.sleep) %>% 
  check_model()
```
## Simulating a linear mixed effects model 

To generate some data for a linear mixed effects model with random intercepts, we do pretty much what we are used to doing when we generated data for a linear model. However, this time, we have an additional parameter that captures the variance in the intercepts between participants. So, we draw a separate (offset from the global) intercept for each participant from this distribution.  

```{r}
# make example reproducible 
set.seed(1)

# parameters
sample_size = 100
b0 = 1
b1 = 2
sd_residual = 1
sd_participant = 0.5 

# randomly draw intercepts for each participant
intercepts = rnorm(sample_size, sd = sd_participant)

# generate the data 
df.mixed = tibble(condition = rep(0:1, each = sample_size), 
                  participant = rep(1:sample_size, 2)) %>% 
  group_by(condition) %>% 
  mutate(value = b0 + b1 * condition + intercepts + rnorm(n(), sd = sd_residual)) %>% 
  ungroup %>% 
  mutate(condition = as.factor(condition),
         participant = as.factor(participant))
```

Let's take a look at the simulated data set: 

```{r}
df.mixed %>% 
  select(participant, condition, value) %>% 
  arrange(participant)
```


Let's fit a model to this data now and take a look at the summary output: 

```{r}
# fit model
fit.mixed = lmer(formula = value ~ 1 + condition + (1 | participant),
                data = df.mixed)

summary(fit.mixed)
```

Let's visualize the model's predictions: 

```{r}
fit.mixed %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```

Let's simulate some data from this fitted model: 

```{r}
# simulated data 
fit.mixed %>% 
  simulate() %>% 
  bind_cols(df.mixed) %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

Even though we only fitted random intercepts in this model, when we simulate from the model, we get different slopes since, when simulating new data, the model takes our uncertainty in the residuals into account as well. 

Let's see whether fitting random intercepts was worth it in this case: 

```{r}
# using chisq test
fit.compact = lm(formula = value ~ 1 +  condition,
                data = df.mixed)

fit.augmented = lmer(formula = value ~ 1 + condition +  (1 | participant),
                data = df.mixed)

anova(fit.augmented, fit.compact)
```

Nope, it's not worth it in this case. That said, even though having random intercepts does not increase the likelihood of the data given the model significantly, we should still include random intercepts to capture the dependence in the data. 

### The effect of outliers 

Let's take 20 participants from our `df.mixed` data set, and make one of the participants be an outlier: 

```{r}
# let's make one outlier
df.outlier = df.mixed %>%
  mutate(participant = participant %>% as.character() %>% as.numeric()) %>% 
  filter(participant <= 20) %>%
  mutate(value = ifelse(participant == 20, value + 30, value),
         participant = as.factor(participant))
```

Let's fit the model and look at the summary: 

```{r}
# fit model
fit.outlier = lmer(formula = value ~ 1 + condition + (1 | participant),
                   data = df.outlier)

summary(fit.outlier)
```

The variance for the participants' intercepts has increased dramatically! 

Let's visualize the data together with the model's predictions: 

```{r}
fit.outlier %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```

The model is still able to capture the participants quite well. But note what its simulated data looks like now: 

```{r}
# simulated data from lmer with outlier
fit.outlier %>% 
  simulate() %>% 
  bind_cols(df.outlier) %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

The simulated data doesn't look like our original data. This is because one normal distribution is used to model the variance in the intercepts between participants. 

### Different slopes 

Let's generate data where the effect of condition is different for participants: 

```{r}
# make example reproducible 
set.seed(1)

tmp = rnorm(n = 20)

df.slopes = tibble(
  condition = rep(1:2, each = 20), 
  participant = rep(1:20, 2),
  value = ifelse(condition == 1, tmp,
                 mean(tmp) + rnorm(n = 20, sd = 0.3)) # regression to the mean
) %>% 
  mutate(condition = as.factor(condition),
         participant = as.factor(participant))
```

Let's fit a model with random intercepts. 

```{r}
fit.slopes = lmer(formula = value ~ 1 + condition + (1 | participant),
                data = df.slopes)

summary(fit.slopes)
```

Note how the summary says "singular fit", and how the variance for random intercepts is 0. Here, fitting random intercepts did not help the model fit at all, so the lmer gave up ... 

How about fitting random slopes? 

```{r, eval=F}
# fit model
lmer(formula = value ~ 1 + condition + (1 + condition | participant),
     data = df.slopes)
```

This won't work because the model has more parameters than there are data points. To fit random slopes, we need more than 2 observations per participants. 

### Simpson's paradox 

Taking dependence in the data into account is extremely important. The Simpson's paradox is an instructive example for what can go wrong when we ignore the dependence in the data. 

Let's start by simulating some data to demonstrate the paradox. 

```{r}
# make example reproducible 
set.seed(2)

n_participants = 20
n_observations = 10
slope = -10 
sd_error = 0.4
sd_participant = 5
intercept = rnorm(n_participants, sd = sd_participant) %>% sort()

df.simpson = tibble(x = runif(n_participants * n_observations, min = 0, max = 1)) %>%
  arrange(x) %>% 
  mutate(intercept = rep(intercept, each = n_observations),
         y = intercept + x * slope + rnorm(n(), sd = sd_error),
         participant = factor(intercept, labels = 1:n_participants))
```

Let's visualize the overall relationship between `x` and `y` with a simple linear model. 

```{r, message=FALSE}
# overall effect 
ggplot(data = df.simpson,
       mapping = aes(x = x,
                     y = y)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "black")
```

As we see, overall, there is a positive relationship between `x` and `y`.

```{r}
lm(formula = y ~ x,
   data = df.simpson) %>% 
  summary()
```

And this relationship is significant. 

Let's take another look at the data use different colors for the different participants.

```{r, message=FALSE}
# effect by participant 
ggplot(data = df.simpson,
       mapping = aes(x = x,
                     y = y,
                     color = participant)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "black") +
  theme(legend.position = "none")
```

And let's fit a different regression for each participant:

```{r, message=FALSE}
# effect by participant 
ggplot(data = df.simpson,
       mapping = aes(x = x,
                     y = y,
                     color = participant,
                     group = participant)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "black") +
  theme(legend.position = "none")
```

What this plot shows, is that for almost all individual participants, the relationship between `x` and `y` is negative. The different participants where along the `x` spectrum they are. 

Let's fit a linear mixed effects model with random intercepts: 

```{r}
fit.lmer = lmer(formula = y ~ 1 + x + (1 | participant),
     data = df.simpson)

fit.lmer %>% 
  summary()
```

As we can see, the fixed effect for `x` is now negative! 

```{r}

fit.lmer %>% 
  augment() %>% 
  clean_names() %>% 
  ggplot(data = .,
         aes(x = x,
             y = y,
             group = participant,
             color = participant)) +
  geom_point() +
  geom_line(aes(y = fitted),
            size = 1,
            color = "black") +
  theme(legend.position = "none")

```

Lesson learned: taking dependence into account is critical for drawing correct inferences! 


## Additional resources 

### Readings 

- [Linear mixed effects models tutorial by Bodo Winter](https://arxiv.org/pdf/1308.5499.pdf)
- [Simpson's paradox](https://paulvanderlaken.com/2017/09/27/simpsons-paradox-two-hr-examples-with-r-code/)
- [Tutorial on pooling](https://www.tjmahr.com/plotting-partial-pooling-in-mixed-effects-models/)

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r}
sessionInfo()
```
