---
title: "22 Bayesian data analysis 2"
author: "Tobias Gerstenberg"
date: "3/6/2019"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

> Little hard to follow today. Having one example (possibly more interesting than coin flips?) to follow throughout the lecture would help scaffold the flow. But I'm still excited about Bayes!

> Concrete examples. Also, some examples from the literature using Bayes rules could be super helpful because we could see what it looks like in action

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-22.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

## What to do next time

- this lecture still needs work 
- it went to fast and was too abstract
- try and make it more concrete (have more opportunity for class engagement)
