---
title: "09_Modeling Data"
author: "Tobias Gerstenberg"
date: "1/27/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

> I didn't fully understand the utility of the f statistic. I like the simple heuristic of modeling as a process of increasing accuracy while minimizing complexity, but I didn't fully understand why it's important to minimize complexity. A short explanation might fix things for me.

> Hope we cover degrees of freedom.

> I think the lecture is still too theoretical with not enough concrete examples....it makes it hard to follow.

> 	I found the modeling-based approach very intuitive and I think it helped me understand the conceptual basis of statistical testing more so than other classes I've taken. Thanks.

## Homework 2

```{r echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, results='hide'}
df.hw_hours = read.csv("how-many-hours-did-it-2.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response,
                           levels = c(1:14, "15 or more"),
                           labels = c(1:14, "15\nor more"))) %>% 
  count(response) %>% 
  complete(response)

ggplot(data = df.hw_hours,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity",
           fill = "lightblue",
           color = "black") +
  scale_y_continuous(breaks = seq(0, 50, 2),
                     expand = c(0, 0)) +
  labs(y = "number of students", x = "", title = "How many hours did you spend on homework 2?")
```

## Student feedback 

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-9.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read.csv("how-was-the-pace-of-9.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-9.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
```

## What to do next time

- this lecture worked quite well 
- didn't manage to cover the last section on how to generate a sampling distribution for PRE 
- try to keep level of interaction high 
- try to make things a little more concrete (if possible)
