---
title: "Class 18"
author: "Tobias Gerstenberg"
date: "February 21st, 2020"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=17"]
bibliography: [packages.bib]
nocite: '@*'
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(comment = "#>",
                      fig.show = "hold")
```

# Linear mixed effects models 2

## Load packages and set plotting theme  

```{r lmer2-01, message=FALSE}
library("knitr")      # for knitting RMarkdown 
library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("patchwork")  # for making figure panels
library("lme4")       # for linear mixed effects models
library("tidyverse")  # for wrangling, plotting, etc. 

# include references for used packages
knitr::write_bib(.packages(), "packages.bib") 
```

```{r lmer2-02}
theme_set(theme_classic() + #set the theme 
            theme(text = element_text(size = 20))) #set the default text size
```

## Simulating a linear mixed effects model 

To generate some data for a linear mixed effects model with random intercepts, we do pretty much what we are used to doing when we generated data for a linear model. However, this time, we have an additional parameter that captures the variance in the intercepts between participants. So, we draw a separate (offset from the global) intercept for each participant from this distribution.  

```{r lmer2-03}
# make example reproducible 
set.seed(1)

# parameters
sample_size = 100
b0 = 1
b1 = 2
sd_residual = 1
sd_participant = 0.5 

# randomly draw intercepts for each participant
intercepts = rnorm(sample_size, sd = sd_participant)

# generate the data 
df.mixed = tibble(condition = rep(0:1, each = sample_size), 
                  participant = rep(1:sample_size, 2)) %>% 
  group_by(condition) %>% 
  mutate(value = b0 + b1 * condition + intercepts + rnorm(n(), sd = sd_residual)) %>% 
  ungroup %>% 
  mutate(condition = as.factor(condition),
         participant = as.factor(participant))
```

Let's take a look at the simulated data set: 

```{r lmer2-04}
df.mixed %>% 
  select(participant, condition, value) %>% 
  arrange(participant) %>% 
  head(8) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
```


Let's fit a model to this data now and take a look at the summary output: 

```{r lmer2-05}
# fit model
fit.mixed = lmer(formula = value ~ 1 + condition + (1 | participant),
                data = df.mixed)

fit.mixed %>% 
  summary()
```

Let's visualize the model's predictions: 

```{r lmer2-06}
fit.mixed %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```

Let's simulate some data from this fitted model: 

```{r lmer2-07}
# simulated data 
fit.mixed %>% 
  simulate() %>% 
  bind_cols(df.mixed) %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

Even though we only fitted random intercepts in this model, when we simulate from the model, we get different slopes since, when simulating new data, the model takes our uncertainty in the residuals into account as well. 

Let's see whether fitting random intercepts was worth it in this case: 

```{r lmer2-08}
# using chisq test
fit.compact = lm(formula = value ~ 1 +  condition,
                data = df.mixed)

fit.augmented = lmer(formula = value ~ 1 + condition +  (1 | participant),
                data = df.mixed)

anova(fit.augmented, fit.compact)
```

Nope, it's not worth it in this case. That said, even though having random intercepts does not increase the likelihood of the data given the model significantly, we should still include random intercepts to capture the dependence in the data. 

## The effect of outliers 

Let's take 20 participants from our `df.mixed` data set, and make one of the participants be an outlier: 

```{r lmer2-09}
# let's make one outlier
df.outlier = df.mixed %>%
  mutate(participant = participant %>% as.character() %>% as.numeric()) %>% 
  filter(participant <= 20) %>%
  mutate(value = ifelse(participant == 20, value + 30, value),
         participant = as.factor(participant))
```

Let's fit the model and look at the summary: 

```{r lmer2-10}
# fit model
fit.outlier = lmer(formula = value ~ 1 + condition + (1 | participant),
                   data = df.outlier)

fit.outlier %>% 
  summary()
```

The variance for the participants' intercepts has increased dramatically! 

Let's visualize the data together with the model's predictions: 

```{r lmer2-11}
fit.outlier %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```

The model is still able to capture the participants quite well. But note what its simulated data looks like now: 

```{r lmer2-12}
# simulated data from lmer with outlier
fit.outlier %>% 
  simulate() %>% 
  bind_cols(df.outlier) %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

The simulated data doesn't look like our original data. This is because one normal distribution is used to model the variance in the intercepts between participants. 

## Different slopes 

Let's generate data where the effect of condition is different for participants: 

```{r lmer2-13}
# make example reproducible 
set.seed(1)

tmp = rnorm(n = 20)

df.slopes = tibble(
  condition = rep(1:2, each = 20), 
  participant = rep(1:20, 2),
  value = ifelse(condition == 1, tmp,
                 mean(tmp) + rnorm(n = 20, sd = 0.3)) # regression to the mean
) %>% 
  mutate(condition = as.factor(condition),
         participant = as.factor(participant))
```

Let's fit a model with random intercepts. 

```{r lmer2-14}
fit.slopes = lmer(formula = value ~ 1 + condition + (1 | participant),
                data = df.slopes)

fit.slopes %>% summary()
```

Note how the summary says "singular fit", and how the variance for random intercepts is 0. Here, fitting random intercepts did not help the model fit at all, so the lmer gave up ... 

How about fitting random slopes? 

```{r lmer2-15, eval=F}
# fit model
lmer(formula = value ~ 1 + condition + (1 + condition | participant),
     data = df.slopes)
```

This won't work because the model has more parameters than there are data points. To fit random slopes, we need more than 2 observations per participants. 

## Simpson's paradox 

Taking dependence in the data into account is extremely important. The Simpson's paradox is an instructive example for what can go wrong when we ignore the dependence in the data. 

Let's start by simulating some data to demonstrate the paradox. 

```{r lmer2-16}
# make example reproducible 
set.seed(2)

n_participants = 20
n_observations = 10
slope = -10 
sd_error = 0.4
sd_participant = 5
intercept = rnorm(n_participants, sd = sd_participant) %>% sort()

df.simpson = tibble(x = runif(n_participants * n_observations, min = 0, max = 1)) %>%
  arrange(x) %>% 
  mutate(intercept = rep(intercept, each = n_observations),
         y = intercept + x * slope + rnorm(n(), sd = sd_error),
         participant = factor(intercept, labels = 1:n_participants))
```

Let's visualize the overall relationship between `x` and `y` with a simple linear model. 

```{r lmer2-17}
# overall effect 
ggplot(data = df.simpson,
       mapping = aes(x = x,
                     y = y)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "black")
```

As we see, overall, there is a positive relationship between `x` and `y`.

```{r lmer2-18}
lm(formula = y ~ x,
   data = df.simpson) %>% 
  summary()
```

And this relationship is significant. 

Let's take another look at the data use different colors for the different participants.

```{r lmer2-19}
# effect by participant 
ggplot(data = df.simpson,
       mapping = aes(x = x,
                     y = y,
                     color = participant)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "black") +
  theme(legend.position = "none")
```

And let's fit a different regression for each participant:

```{r lmer2-20}
# effect by participant 
ggplot(data = df.simpson,
       mapping = aes(x = x,
                     y = y,
                     color = participant,
                     group = participant)) +
  geom_point() +
  geom_smooth(method = "lm",
              color = "black") +
  theme(legend.position = "none")
```

What this plot shows, is that for almost all individual participants, the relationship between `x` and `y` is negative. The different participants where along the `x` spectrum they are. 

Let's fit a linear mixed effects model with random intercepts: 

```{r lmer2-21}
fit.lmer = lmer(formula = y ~ 1 + x + (1 | participant),
     data = df.simpson)

fit.lmer %>% 
  summary()
```

As we can see, the fixed effect for `x` is now negative! 

```{r lmer2-22}

fit.lmer %>% 
  augment() %>% 
  clean_names() %>% 
  ggplot(data = .,
         aes(x = x,
             y = y,
             group = participant,
             color = participant)) +
  geom_point() +
  geom_line(aes(y = fitted),
            size = 1,
            color = "black") +
  theme(legend.position = "none")

```

Lesson learned: taking dependence into account is critical for drawing correct inferences! 

## Additional resources 

### Readings 

- [Linear mixed effects models tutorial by Bodo Winter](https://arxiv.org/pdf/1308.5499.pdf)
- [Simpson's paradox](https://paulvanderlaken.com/2017/09/27/simpsons-paradox-two-hr-examples-with-r-code/)
- [Tutorial on pooling](https://www.tjmahr.com/plotting-partial-pooling-in-mixed-effects-models/)

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r lmer2-23}
sessionInfo()
```

## References