---
title: "Class 13"
author: "Tobias Gerstenberg"
date: "February 10th, 2021"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=12"]
bibliography: [packages.bib]
nocite: '@*'
---

# Linear model 4

```{r, include=FALSE, eval=FALSE}
install.packages("afex")
```

## Load packages and set plotting theme  

```{r load-packages, message=FALSE}
library("knitr")      # for knitting RMarkdown 
library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("afex")       # for running ANOVAs
library("emmeans")    # for calculating contrasts
library("car")        # for calculating ANOVAs
library("tidyverse")  # for wrangling, plotting, etc.
```

```{r}
theme_set(
  theme_classic() + #set the theme 
    theme(text = element_text(size = 20)) #set the default text size
)

# these options here change the formatting of how comments are rendered
opts_chunk$set(comment = "",
               fig.show = "hold")

# include references for used packages
write_bib(.packages(), "packages.bib") 
```

## Load data sets

Read in the data:

```{r, message=FALSE}
df.poker = read_csv("data/poker.csv") %>% 
  mutate(skill = factor(skill,
                        levels = 1:2,
                        labels = c("expert", "average")),
         skill = fct_relevel(skill, "average", "expert"),
         hand = factor(hand,
                       levels = 1:3,
                       labels = c("bad", "neutral", "good")),
         limit = factor(limit,
                        levels = 1:2,
                        labels = c("fixed", "none")),
         participant = 1:n()) %>% 
  select(participant, everything())
```

## Linear contrasts 

Here is a linear contrast that assumes that there is a linear relationship between the quality of one's hand, and the final balance.  

```{r}
df.poker = df.poker %>% 
  mutate(hand_contrast = factor(hand,
                                levels = c("bad", "neutral", "good"),
                                labels = c(-1, 0, 1)),
         hand_contrast = hand_contrast %>% as.character() %>% as.numeric())

fit.contrast = lm(formula = balance ~ hand_contrast,
                  data = df.poker)
```

Here is a visualization of the model prediction together with the residuals. 

```{r}
df.plot = df.poker %>% 
  mutate(hand_jitter = hand %>% as.numeric(),
         hand_jitter = hand_jitter + runif(n(), min = -0.4, max = 0.4))

df.tidy = fit.contrast %>% 
  tidy() %>% 
  select_if(is.numeric) %>% 
  mutate_all(~ round(., 2))

df.augment = fit.contrast %>% 
  augment() %>%
  clean_names() %>% 
  bind_cols(df.plot %>% select(hand_jitter))

ggplot(data = df.plot,
       mapping = aes(x = hand_jitter,
                       y = balance,
                       color = as.factor(hand_contrast))) + 
  geom_point(alpha = 0.8) +
  geom_segment(data = NULL,
               aes(x = 0.6,
                   xend = 1.4,
                   y = df.tidy$estimate[1]-df.tidy$estimate[2],
                   yend = df.tidy$estimate[1]-df.tidy$estimate[2]),
               color = "red",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1],
                   yend = df.tidy$estimate[1]),
               color = "orange",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "green",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = hand_jitter,
                   y = balance,
                   yend = fitted),
               alpha = 0.3) +
  labs(y = "balance") + 
  scale_color_manual(values = c("red", "orange", "green")) + 
  scale_x_continuous(breaks = 1:3, labels = c("bad", "neutral", "good")) + 
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

### Hypothetical data 

Here is some code to generate a hypothetical developmental data set. 

```{r}
# make example reproducible 
set.seed(1)

# means = c(5, 10, 5)
means = c(3, 5, 20)
# means = c(3, 5, 7)
# means = c(3, 7, 12)
sd = 2
sample_size = 20

# generate data 
df.development = tibble(
  group = rep(c("3-4", "5-6", "7-8"), each = sample_size),
  performance = NA) %>% 
  mutate(performance = ifelse(group == "3-4",
                              rnorm(sample_size,
                                    mean = means[1],
                                    sd = sd),
                              performance),
         performance = ifelse(group == "5-6",
                              rnorm(sample_size,
                                    mean = means[2],
                                    sd = sd),
                              performance),
         performance = ifelse(group == "7-8",
                              rnorm(sample_size,
                                    mean = means[3],
                                    sd = sd),
                              performance),
         group = factor(group, levels = c("3-4", "5-6", "7-8")),
         group_contrast = group %>% 
           fct_recode(`-1` = "3-4",
                      `0` = "5-6",
                      `1` = "7-8") %>% 
           as.character() %>%
           as.numeric())
```

Let's define a linear contrast using the `emmeans` package, and test whether it's significant. 

```{r}
fit = lm(formula = performance ~ group,
         data = df.development)

fit %>% 
  emmeans("group",
          contr = list(linear = c(-0.5, 0, 0.5)),
          adjust = "bonferroni") %>% 
  pluck("contrasts")
```

Yes, we see that there is a significant positive linear contrast with an estimate of 8.45. This means, it predicts a difference of 8.45 in performance between each of the consecutive age groups. For a visualization of the predictions of this model, see Figure \@ref{fig:linear-contrast-model}. 

### Visualization

Total variance: 

```{r}
set.seed(1)

fit_c = lm(formula = performance ~ 1,
           data = df.development)

df.plot = df.development %>% 
  mutate(group_jitter = 1 + runif(n(), min = -0.25, max = 0.25))

df.augment = fit_c %>% 
  augment() %>% 
  clean_names() %>% 
  bind_cols(df.plot %>% select(group, group_jitter))

ggplot(data = df.plot, 
       mapping = aes(x = group_jitter,
                       y = performance,
                       fill = group)) + 
  geom_hline(yintercept = mean(df.development$performance)) +
  geom_point(alpha = 0.5) + 
  geom_segment(data = df.augment,
               aes(xend = group_jitter,
                   yend = fitted),
               alpha = 0.2) +
  labs(y = "performance") + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank())

```

With contrast

```{r linear-contrast-model, fig.cap="Predictions of the linear contrast model"}
# make example reproducible 
set.seed(1)

fit = lm(formula = performance ~ group_contrast,
         data = df.development)

df.plot = df.development %>% 
  mutate(group_jitter = group %>% as.numeric(),
         group_jitter = group_jitter + runif(n(), min = -0.4, max = 0.4))

df.tidy = fit %>% 
  tidy() %>% 
  select(where(is.numeric)) %>% 
  mutate(across(.fns = ~ round(. , 2)))

df.augment = fit %>% 
  augment() %>%
  clean_names() %>% 
  bind_cols(df.plot %>% select(group_jitter))

ggplot(data = df.plot,
       mapping = aes(x = group_jitter,
                       y = performance,
                       color = as.factor(group_contrast))) + 
  geom_point(alpha = 0.8) +
  geom_segment(data = NULL,
               aes(x = 0.6,
                   xend = 1.4,
                   y = df.tidy$estimate[1]-df.tidy$estimate[2],
                   yend = df.tidy$estimate[1]-df.tidy$estimate[2]),
               color = "red",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 1.6,
                   xend = 2.4,
                   y = df.tidy$estimate[1],
                   yend = df.tidy$estimate[1]),
               color = "orange",
               size = 1) +
  geom_segment(data = NULL,
               aes(x = 2.6,
                   xend = 3.4,
                   y = df.tidy$estimate[1] + df.tidy$estimate[2],
                   yend = df.tidy$estimate[1] + df.tidy$estimate[2]),
               color = "green",
               size = 1) +
  geom_segment(data = df.augment,
               aes(xend = group_jitter,
                   y = performance,
                   yend = fitted),
               alpha = 0.3) +
  labs(y = "performance") + 
  scale_color_manual(values = c("red", "orange", "green")) + 
  scale_x_continuous(breaks = 1:3, labels = levels(df.development$group)) +
  theme(legend.position = "none",
        axis.title.x = element_blank())
```

Results figure

```{r}
df.development %>% 
  ggplot(aes(x = group, y = performance)) + 
  geom_point(alpha = 0.3, position = position_jitter(width = 0.1, height = 0)) +
  stat_summary(fun.data = "mean_cl_boot",
               shape = 21, 
               fill = "white",
               size = 0.75)
```

Here we test some more specific hypotheses: the the two youngest groups of children are different from the oldest group, and that the 3 year olds are different from the 5 year olds. 

```{r}
#  fit the linear model 
fit = lm(formula = performance ~ group,
         data = df.development)

# check factor levels 
levels(df.development$group)

# define the contrasts of interest 
contrasts = list(young_vs_old = c(-0.5, -0.5, 1),
                 three_vs_five = c(-0.5, 0.5, 0))

# compute significance test on contrasts 
fit %>% 
  emmeans("group",
          contr = contrasts,
          adjust = "bonferroni") %>% 
  pluck("contrasts")
```

### Post-hoc tests

Post-hoc tests for a single predictor (using the poker data set). 

```{r}
fit = lm(formula = balance ~ hand,
         data = df.poker)

# post hoc tests 
fit %>% 
  emmeans(pairwise ~ hand,
          adjust = "bonferroni") %>% 
  pluck("contrasts")
```

Post-hoc tests for two predictors (:

```{r}
# fit the model
fit = lm(formula = balance ~ hand + skill,
         data = df.poker)

# post hoc tests 
fit %>% 
  emmeans(pairwise ~ hand + skill,
          adjust = "bonferroni") %>% 
  pluck("contrasts")
```


```{r}
fit = lm(formula = balance ~ hand,
         data = df.poker)

# comparing each to the mean 
fit %>% 
  emmeans(eff ~ hand) %>% 
  pluck("contrasts")

# one vs. all others 
fit %>% 
  emmeans(del.eff ~ hand) %>% 
  pluck("contrasts")
```

### Understanding dummy coding 

```{r}
fit = lm(formula = balance ~ 1 + hand,
         data = df.poker)

fit %>% 
  summary()

model.matrix(fit) %>% 
  as_tibble() %>% 
  distinct()

df.poker %>% 
  select(participant, hand, balance) %>% 
  group_by(hand) %>% 
  top_n(3, wt = -participant) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
```

### Understanding effect coding 

```{r}
fit = lm(formula = balance ~ 1 + hand,
         contrasts = list(hand = "contr.sum"),
         data = df.poker)

fit %>% 
  summary()

model.matrix(fit) %>% 
  as_tibble() %>% 
  distinct() %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
```

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r session}
sessionInfo()
```

## References