---
title: "Class 17"
author: "Tobias Gerstenberg"
date: "February 20th, 2019"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=16"]
bibliography: [packages.bib]
nocite: '@*'
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(comment = "#>",
                      fig.show = "hold")
```

# Linear mixed effects models 1

```{r lmer1-01, include=FALSE, eval=FALSE}
install.packages(c("lme4", "lmerTest", "pbkrtest"))
```

## Load packages and set plotting theme  

```{r lmer1-02, message=FALSE}
library("knitr")      # for knitting RMarkdown 
library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("patchwork")  # for making figure panels
library("lme4")       # for linear mixed effects models
library("tidyverse")  # for wrangling, plotting, etc. 

# include references for used packages
knitr::write_bib(.packages(), "packages.bib") 
```

```{r lmer1-03}
theme_set(theme_classic() + #set the theme 
            theme(text = element_text(size = 20))) #set the default text size
```

## Dependence 

Let's generate a data set in which two observations from the same participants are dependent, and then let's also shuffle this data set to see whether taking into account the dependence in the data matters. 

```{r lmer1-04}
# make example reproducible 
set.seed(1)

df.dependence = tibble(participant = 1:20,
                       condition1 = rnorm(20),
                       condition2 = condition1 + rnorm(20, mean = 0.2, sd = 0.1)) %>% 
  mutate(condition2shuffled = sample(condition2)) # shuffles the condition label
```

Let's visualize the original and shuffled data set: 

```{r lmer1-05}
df.plot = df.dependence %>% 
  pivot_longer(cols = -participant,
               names_to = "condition",
               values_to = "value") %>% 
  mutate(condition = str_replace(condition, "condition", ""))

p1 = ggplot(data = df.plot %>% 
              filter(condition != "2shuffled"), 
            mapping = aes(x = condition, y = value)) +
  geom_line(aes(group = participant), alpha = 0.3) +
  geom_point() +
  stat_summary(fun.y = "mean", 
               geom = "point",
               shape = 21, 
               fill = "red",
               size = 4) +
  labs(title = "original",
       tag = "a)")

p2 = ggplot(data = df.plot %>% 
              filter(condition != "2"), 
            mapping = aes(x = condition, y = value)) +
  geom_line(aes(group = participant), alpha = 0.3) +
  geom_point() +
  stat_summary(fun.y = "mean", 
               geom = "point",
               shape = 21, 
               fill = "red",
               size = 4) +
  labs(title = "shuffled",
       tag = "b)")

p1 + p2 
```

Let's save the two original and shuffled data set as two separate data sets.

```{r lmer1-06}
# separate the data sets 
df.original = df.dependence %>% 
  pivot_longer(cols = -participant,
               names_to = "condition",
               values_to = "value") %>% 
  mutate(condition = str_replace(condition, "condition", "")) %>% 
  filter(condition != "2shuffled")

df.shuffled = df.dependence %>% 
  pivot_longer(cols = -participant,
               names_to = "condition",
               values_to = "value") %>% 
  mutate(condition = str_replace(condition, "condition", "")) %>% 
  filter(condition != "2")
```

Let's run a linear model, and independent samples t-test on the original data set. 

```{r lmer1-07}
# linear model (assuming independent samples)
lm(formula = value ~ condition,
   data = df.original) %>% 
  summary() 

t.test(df.original$value[df.original$condition == "1"],
       df.original$value[df.original$condition == "2"],
       alternative = "two.sided",
       paired = F)
```

The mean difference between the conditions is extremely small, and non-significant (if we ignore the dependence in the data). 

Let's fit a linear mixed effects model with a random intercept for each participant: 

```{r lmer1-08}
# fit a linear mixed effects model 
lmer(formula = value ~ condition + (1 | participant),
     data = df.original) %>% 
  summary()
```

To test for whether condition is a significant predictor, we need to use our model comparison approach: 

```{r lmer1-09}
# fit models
fit.compact = lmer(formula = value ~ 1 + (1 | participant),
                   data = df.original)
fit.augmented = lmer(formula = value ~ condition + (1 | participant),
                     data = df.original)

# compare via Chisq-test
anova(fit.compact, fit.augmented)
```

This result is identical to running a paired samples t-test: 

```{r lmer1-10}
t.test(df.original$value[df.original$condition == "1"],
       df.original$value[df.original$condition == "2"],
       alternative = "two.sided",
       paired = T)
```

But, unlike in the paired samples t-test, the linear mixed effects model explicitly models the variation between participants, and it's a much more flexible approach for modeling dependence in data. 

Let's fit a linear model and a linear mixed effects model to the original (non-shuffled) data. 

```{r lmer1-11}
# model assuming independence
fit.independent = lm(formula = value ~ 1 + condition,
                     data = df.original)

# model assuming dependence
fit.dependent = lmer(formula = value ~ 1 + condition + (1 | participant),
                     data = df.original)
```

Let's visualize the linear model's predictions: 

```{r lmer1-12}
# plot with predictions by fit.independent 
fit.independent %>% 
  augment() %>% 
  bind_cols(df.original %>%
              select(participant)) %>% 
  clean_names() %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") + 
  geom_line(aes(y = fitted),
            color = "red")
```

And this is what the residuals look like: 

```{r lmer1-13}
# make example reproducible 
set.seed(1)

fit.independent %>% 
  augment() %>% 
  bind_cols(df.original %>%
              select(participant)) %>% 
  clean_names() %>% 
  mutate(index = as.numeric(condition),
         index = index + runif(n(), min = -0.3, max = 0.3)) %>% 
  ggplot(data = .,
         mapping = aes(x = index,
                       y = value,
                       group = participant,
                       color = condition)) +
  geom_point() + 
  geom_smooth(method = "lm",
              se = F,
              formula = "y ~ 1",
              aes(group = condition)) +
  geom_segment(aes(xend = index,
                   yend = fitted),
               alpha = 0.5) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(breaks = 1:2, 
                     labels = 1:2) +
  labs(x = "condition") +
  theme(legend.position = "none")

```

It's clear from this residual plot, that fitting two separate lines (or points) is not much better than just fitting one line (or point). 

Let's visualize the predictions of the linear mixed effects model: 

```{r lmer1-14}
# plot with predictions by fit.independent 
fit.dependent %>% 
  augment() %>% 
  clean_names() %>% 
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") + 
  geom_line(aes(y = fitted),
            color = "red")
```

Let's compare the residuals of the linear model with that of the linear mixed effects model: 

```{r lmer1-15}
# linear model 
p1 = fit.independent %>% 
  augment() %>% 
  clean_names() %>% 
  ggplot(data = .,
         mapping = aes(x = fitted,
                       y = resid)) +
  geom_point() +
  coord_cartesian(ylim = c(-2.5, 2.5))

# linear mixed effects model 
p2 = fit.dependent %>% 
  augment() %>% 
  clean_names() %>% 
  ggplot(data = .,
         mapping = aes(x = fitted,
                       y = resid)) +
  geom_point() + 
  coord_cartesian(ylim = c(-2.5, 2.5))

p1 + p2
```

The residuals of the linear mixed effects model are much smaller. Let's test whether taking the individual variation into account is worth it (statistically speaking). 

```{r lmer1-16}
# fit models (without and with dependence)
fit.compact = lm(formula = value ~ 1 + condition,
                 data = df.original)

fit.augmented = lmer(formula = value ~ 1 + condition + (1 | participant),
                     data = df.original)

# compare models
# note: the lmer model has to be supplied first 
anova(fit.augmented, fit.compact) 
```

Yes, the likelihood of the data given the linear mixed effects model is significantly higher compared to its likelihood given the linear model. 

## Additional resources 

### Readings 

- [Linear mixed effects models tutorial by Bodo Winter](https://arxiv.org/pdf/1308.5499.pdf)

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r lmer1-17}
sessionInfo()
```

## References