---
title: "Class 6"
author: "Tobias Gerstenberg"
date: "January 25th, 2020"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=5"]
---

# Probability and causality

## Install and load packages, load data, set theme

Let's first install the new packages that you might not have yet. 

```{r, eval=FALSE}
install.packages(c("arrangements", "kableExtra", "DiagrammeR"))
```

Let's load the packages that we need for this chapter. 

```{r, message=FALSE}
library("knitr")        # for rendering the RMarkdown file
library("kableExtra")   # for nicely formatted tables
library("arrangements") # fast generators and iterators for creating combinations
library("DiagrammeR")   # for drawing diagrams
library("tidyverse")    # for data wrangling 
```

Set the plotting theme.

```{r}
theme_set(theme_classic() + 
            theme(text = element_text(size = 20)))

opts_chunk$set(comment = "",
               fig.show = "hold")
```


## Counting 

Imagine that there are three balls in an urn. The balls are labeled 1, 2, and 3. Let's consider a few possible situations. 

```{r}
balls = 1:3 # number of balls in urn 
ndraws = 2 # number of draws

# order matters, without replacement
permutations(balls, ndraws)

# order matters, with replacement
permutations(balls, ndraws, replace = T)

# order doesn't matter, with replacement 
combinations(balls, ndraws, replace = T)

# order doesn't matter, without replacement 
combinations(balls, ndraws)
```

I've generated the figures below using the `DiagrammeR` package. It's a powerful package for drawing diagrams in R. See information on how to use the DiagrammeR package [here](http://rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html). 

```{r, echo=FALSE, fig.cap="Drawing two marbles out of an urn __with__ replacement."}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 24,
    fillcolor = lightblue
    ]
  
  # edges between nodes
  edge [color = black]
  0 -> {1 2 3}
  1 -> {11 12 13}
  2 -> {21 22 23}
  3 -> {31 32 33}
  
  # labels for each node
  0 [fillcolor = 'black', width = 0.1]
  1 [label = '1']
  2 [label = '2']
  3 [label = '3']
  11 [label = '1']
  12 [label = '2']
  13 [label = '3']
  21 [label = '1']
  22 [label = '2']
  23 [label = '3']
  31 [label = '1']
  32 [label = '2']
  33 [label = '3']
    
  # direction in which arrows are drawn (from left to right)
  rankdir = LR
}
")

```

```{r, echo=FALSE, fig.cap="Drawing two marbles out of an urn __without__ replacement."}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 24,
    fillcolor = lightblue
    ]
  
  # edges between nodes
  edge [color = black]
  0 -> {1 2 3}
  1 -> {12 13}
  2 -> {21 23}
  3 -> {31 32}
  
  # labels for each node
  0 [fillcolor = 'black', width = 0.1]
  1 [label = '1']
  2 [label = '2']
  3 [label = '3']
  12 [label = '2']
  13 [label = '3']
  21 [label = '1']
  23 [label = '3']
  31 [label = '1']
  32 [label = '2']
  
  # direction in which arrows are drawn (from left to right)
  rankdir = LR
}
")
```

## The random secretary

A secretary types four letters to four people and addresses the four envelopes. If he inserts the letters at random, each in a different envelope, what is the probability that exactly three letters will go into the right envelope?

```{r}
df.letters = permutations(x = 1:4, k = 4) %>% 
  as_tibble(.name_repair = ~ str_c("person_", 1:4)) %>%
  mutate(n_correct = (person_1 == 1) + 
           (person_2 == 2) + 
           (person_3 == 3) +
           (person_4 == 4))

df.letters %>% 
  summarize(prob_3_correct = sum(n_correct == 3) / n())
```

```{r}
ggplot(data = df.letters,
       mapping = aes(x = n_correct)) + 
  geom_bar(aes(y = stat(count)/sum(count)),
           color = "black",
           fill = "lightblue") +
  scale_y_continuous(labels = scales::percent,
                     expand = c(0, 0)) + 
  labs(x = "number correct",
       y = "probability")
```

## Flipping a coin many times 

```{r, fig.cap='A demonstration of the law of large numbers.'}

# Example taken from here: http://statsthinking21.org/probability.html#empirical-frequency

set.seed(1) # set the seed so that the outcome is consistent
nsamples = 50000 # how many flips do we want to make?

# create some random coin flips using the rbinom() function with
# a true probability of 0.5

df.samples = tibble(trial_number = seq(nsamples), 
                    outcomes = rbinom(nsamples, 1, 0.5)) %>% 
  mutate(mean_probability = cumsum(outcomes) / seq_along(outcomes)) %>% 
  filter(trial_number >= 10) # start with a minimum sample of 10 flips

ggplot(data = df.samples, 
       mapping = aes(x = trial_number, y = mean_probability)) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed") +
  geom_line() +
  labs(x = "Number of trials",
       y = "Estimated probability of heads") +
  theme_classic() +
  theme(text = element_text(size = 20))
```

## Clue guide to probability 

```{r}
who = c("ms_scarlet", "col_mustard", "mrs_white",
        "mr_green", "mrs_peacock", "prof_plum")
what = c("candlestick", "knife", "lead_pipe",
         "revolver", "rope", "wrench")
where = c("study", "kitchen", "conservatory",
          "lounge", "billiard_room", "hall",
          "dining_room", "ballroom", "library")

df.clue = expand_grid(who = who,
                      what = what,
                      where = where)

df.suspects = df.clue %>% 
  distinct(who) %>% 
  mutate(gender = ifelse(test = who %in% c("ms_scarlet", "mrs_white", "mrs_peacock"), 
                         yes = "female", 
                         no = "male"))
```

```{r}
df.suspects %>% 
  arrange(desc(gender)) %>% 
  kable() %>% 
  kable_styling("striped", full_width = F)
```

### Conditional probability 

```{r}
# conditional probability (via rules of probability)
df.suspects %>% 
  summarize(p_prof_plum_given_male = 
              sum(gender == "male" & who == "prof_plum") /
              sum(gender == "male"))
```
```{r}
# conditional probability (via rejection)
df.suspects %>% 
  filter(gender == "male") %>% 
  summarize(p_prof_plum_given_male = 
              sum(who == "prof_plum") /
              n())
```

### Law of total probability

```{r, echo=FALSE}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 9,
    fillcolor = lightblue,
    fixedsize=true,
    width = 0.8
    ]
  
  # edges between nodes
  edge [color = black,
        fontname = 'Helvetica',
        fontsize = 10]
  1 -> 2 [label = 'p(female)']
  1 -> 3 [label = 'p(male)']
  2 -> 4 [label = 'p(revolver | female)'] 
  3 -> 4 [label = 'p(revolver | male)']
  
  

  # labels for each node
  1 [label = 'Gender?']
  2 [label = 'If female\nuse revolver?']
  3 [label = 'If male\nuse revolver?']
  4 [label = 'Revolver\nused?']
  
  rankdir='LR'
  }"
)
```

## Probability operations 

```{r}
# Make a deck of cards 
df.cards = tibble(suit = rep(c("Clubs", "Spades", "Hearts", "Diamonds"), each = 8),
                  value = rep(c("7", "8", "9", "10", "Jack", "Queen", "King", "Ace"), 4)) 
```

```{r}
# conditional probability: p(Hearts | Queen) (via rules of probability)
df.cards %>% 
  summarize(p_hearts_given_queen = 
              sum(suit == "Hearts" & value == "Queen") / 
              sum(value == "Queen"))
```

```{r}
# conditional probability: p(Hearts | Queen) (via rejection)
df.cards %>% 
  filter(value == "Queen") %>%
  summarize(p_hearts_given_queen = sum(suit == "Hearts")/n())
```

## Bayesian reasoning example

```{r, echo=FALSE}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 10,
    fillcolor = lightblue,
    fixedsize=true,
    width = 0.8
    ]
  
  # edges between nodes
  edge [color = black,
        fontname = 'Helvetica',
        fontsize = 10]
  1 -> 2 [label = 'ill']
  1 -> 3 [label = 'healthy']
  2 -> 4 [label = 'test +'] 
  2 -> 5 [label = 'test -']
  3 -> 6 [label = 'test +']
  3 -> 7 [label = 'test -']
  

  # labels for each node
  1 [label = '10000\npeople']
  2 [label = '100']
  3 [label = '9900']
  4 [label = '95']
  5 [label = '5']
  6 [label = '495']
  7 [label = '9405']
  
  rankdir='LR'
  }"
)
```

## Bayesian networks 

### Sprinkler example

```{r, echo=FALSE}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 10,
    fillcolor = lightblue,
    fixedsize=true,
    width = 0.8
    ]
  
  # edges between nodes
  edge [color = black,
        fontname = 'Helvetica',
        fontsize = 10]
  1 -> 2 [label = '']
  1 -> 3 [label = '']
  2 -> 4 [label = ''] 
  3 -> 4 [label = '']
  
  # labels for each node
  1 [label = 'Cloudy']
  2 [label = 'Sprinkler']
  3 [label = 'Rain']
  4 [label = 'Wet grass']
  }"
)
```

```{r}
# cloudy 
df.cloudy = tibble(`p(C)` = 0.5)

df.cloudy %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F,
                font_size = 20)

# sprinkler given cloudy 
df.sprinkler_given_cloudy = tibble(C = c("F", "T"),
                                   `p(S)`= c(0.5, 0.1))

df.sprinkler_given_cloudy %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F,
                font_size = 20)

# rain given cloudy 
df.rain_given_cloudy = tibble(C = c("F", "T"),
                              `p(R)`= c(0.2, 0.8))

df.rain_given_cloudy %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F,
                font_size = 20)

# wet given sprinkler and rain  
df.rain_given_sprinkler_and_rain = tibble(
  S = rep(c("F", "T"), 2),
  R = rep(c("F", "T"), each = 2),
  `p(W)`= c(0, 0.9, 0.9, 0.99)
)

df.rain_given_sprinkler_and_rain %>% 
  kable() %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F,
                font_size = 20)
```

## Additional resources 

### Cheatsheets 

- [Probability cheatsheet](figures/probability.pdf)

### Books and chapters

- [Probability and Statistics with examples using R](http://www.isibang.ac.in/~athreya/psweur/)
- [Learning statistics with R: Chapter 9 Introduction to probability](https://learningstatisticswithr-bookdown.netlify.com/probability.html#probstats)

### Misc 

- [Bayes' theorem in three panels](https://www.tjmahr.com/bayes-theorem-in-three-panels/)
- [Statistics 110: Probability; course at Harvard](https://projects.iq.harvard.edu/stat110)  
- [Bayes theorem and making probability intuitive](https://www.youtube.com/watch?v=HZGCoVF3YvM&feature=youtu.be)
- [Causal inference with Bayes rule](https://gradientinstitute.org/blog/6/)
  
## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r, echo=F}
sessionInfo()
```
