---
title: "Class 6"
author: "Tobias Gerstenberg"
date: "January 18th, 2019"
output: html_notebook
bibliography: ../references.bib
biblio-style: apalike
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

# Probability

## Learning objectives 

- 

## Load packages 

Let's first load the packages that we need for this chapter. 

```{r, message=FALSE}
library("knitr") # for rendering the RMarkdown file
library("arrangements") # fast generators and iterators for permutations, combinations and partitions
library("DiagrammeR") # for drawing diagrams
library("tidyverse") # for data wrangling 
```

## Counting 

Imagine that there are three balls in an urn. The balls are labeled 1, 2, and 3. Let's consider a few possible situations. 


```{r}
balls = 1:3 # number of balls in urn 
ndraws = 2 # number of draws

# order doesn't matter, with replacement 
combinations(balls, ndraws, replace = T)

# order doesn't matter, without replacement 
combinations(balls, ndraws)

# order matters, without replacement
permutations(balls, ndraws)

# order matters, with replacement
permutations(balls, ndraws, replace = T)
```

```{r, fig.cap="Drawing two marbles out of an urn __with__ replacement."}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 24,
    fillcolor = lightblue
    ]
  
  # edges between nodes
  edge [color = black]
  0 -> {1 2 3}
  1 -> {11 12 13}
  2 -> {21 22 23}
  3 -> {31 32 33}
  
  # labels for each node
  0 [fillcolor = 'black', width = 0.1]
  1 [label = '1']
  2 [label = '2']
  3 [label = '3']
  11 [label = '1']
  12 [label = '2']
  13 [label = '3']
  21 [label = '1']
  22 [label = '2']
  23 [label = '3']
  31 [label = '1']
  32 [label = '2']
  33 [label = '3']
    
  # direction in which arrows are drawn (from left to right)
  rankdir = LR
}
")

```
See information on how to use the DiagrammeR package [here](http://rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html). 

```{r, fig.cap="Drawing two marbles out of an urn __without__ replacement."}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 24,
    fillcolor = lightblue
    ]
  
  # edges between nodes
  edge [color = black]
  0 -> {1 2 3}
  1 -> {12 13}
  2 -> {21 23}
  3 -> {31 32}
  
  # labels for each node
  0 [fillcolor = 'black', width = 0.1]
  1 [label = '1']
  2 [label = '2']
  3 [label = '3']
  12 [label = '2']
  13 [label = '3']
  21 [label = '1']
  23 [label = '3']
  31 [label = '1']
  32 [label = '2']
  
  # direction in which arrows are drawn (from left to right)
  rankdir = LR
}
")
```

## Flipping a coin many times 

```{r FlipSim,echo=FALSE,fig.cap='A demonstration of the law of large numbers.'}

# Example taken from here: http://statsthinking21.org/probability.html#empirical-frequency

set.seed(1) # set the seed so that the outcome is consistent
nsamples = 50000 # how many flips do we want to make?

# create some random coin flips using the rbinom() function with
# a true probability of 0.5

df.samples = tibble(
  trial_number = seq(nsamples), 
  outcomes = rbinom(nsamples, 1, 0.5)) %>% 
  mutate(mean_probability = cumsum(outcomes) / seq_along(outcomes)) %>% 
  filter(trial_number >= 10) # start with a minimum sample of 10 flips

ggplot(data = df.samples, 
       mapping = aes(x = trial_number, y = mean_probability)) +
  geom_hline(yintercept = 0.5, color = "gray", linetype = "dashed") +
  geom_line() +
  labs(
    x = "Number of trials",
    y = "Estimated probability of heads"
  )+
  theme_classic()+
  theme(text = element_text(size = 20))
# ggsave("figures/flipping_coins.png", width = 12, height = 6)
```

## Probability operations 

```{r}
# Make a deck of cards 
df.cards = tibble(
  suit = rep(c("Clubs", "Spades", "Hearts", "Diamonds"), each = 8),
  value = rep(c("7", "8", "9", "10", "Jack", "Queen", "King", "Ace"), 4)
) 
```

```{r}
# conditional probability: p(Hearts | Queen) (via rules of probability)
df.cards %>% 
  summarize(p_hearts_given_queen = 
              sum(suit == "Hearts" & value == "Queen") / 
              sum(value == "Queen"))
```

```{r}
# conditional probability: p(Hearts | Queen) (via rejection)
df.cards %>% 
  filter(value == "Queen") %>%
  summarize(p_hearts_given_queen = sum(suit == "Hearts")/n())
```


## Bayesian updating

- Example from 

```{r}
grViz("
digraph dot{
  
  # general settings for all nodes
  node [
    shape = circle,
    style = filled,
    color = black,
    label = ''
    fontname = 'Helvetica',
    fontsize = 10,
    fillcolor = lightblue,
    fixedsize=true,
    width = 0.8
    ]
  
  # edges between nodes
  edge [color = black,
        fontname = 'Helvetica',
        fontsize = 10]
  1 -> 2 [label = 'ill']
  1 -> 3 [label = 'healthy']
  2 -> 4 [label = 'test +'] 
  2 -> 5 [label = 'test -']
  3 -> 6 [label = 'test +']
  3 -> 7 [label = 'test -']
  

  # labels for each node
  1 [label = '10000\npeople']
  2 [label = '100']
  3 [label = '9900']
  4 [label = '95']
  5 [label = '5']
  6 [label = '495']
  7 [label = '9405']
  
  rankdir='LR'
  }"
)
```

## Concepts to teach 

- Counting 
  - naive definition of probability 
  - formal definition of probability 

- Thinking conditionally 
  - Simpson's paradox 
  - law of total probability 
  - Bayes' rule 

- Causal Bayesian Networks 
  - 

- Law of large numbers 
  - Central limit theorem

- Probability distributions 
  - continuous 
  - discrete
  

## Additional resources 

### Cheatsheets 

- [Probability cheatsheet](figures/probability.pdf)

### Data camp courses 

- [Foundations of probability in R](https://www.datacamp.com/courses/foundations-of-probability-in-r)

### Books and chapters

- [Probability and Statistics with examples using R](http://www.isibang.ac.in/~athreya/psweur/)
- [Learning statistics with R: Chapter 9 Introduction to probability](https://learningstatisticswithr-bookdown.netlify.com/probability.html#probstats)

### Tutorials 

### Misc 

- [Statistics 110: Probability; course at Harvard](https://projects.iq.harvard.edu/stat110)  
  
## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r}
sessionInfo()
```
