---
title: "Class 6"
author: "Tobias Gerstenberg"
date: "January 16th, 2019"
output:
  # html_document
  bookdown::gitbook:
    config:
      toc:
        collapse: none
        scroll_highlight: yes
        before: null
        after: null
      toolbar:
        position: fixed
      edit : null
      search: yes
      fontsettings:
        theme: white
        family: sans
        size: 2
      sharing: no
    toc_depth: 3
bibliography: ../references.bib
biblio-style: apalike
css: ../style.css
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
```

# Probability

## Learning objectives 

- 

## Load packages 

Let's first load the packages that we need for this chapter. 

```{r, message=FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
```

## Concepts to teach 

- Counting 
  - naive definition of probability 
  - formal definition of probability 

- Thinking conditionally 
  - Simpson's paradox 
  - law of total probability 
  - Bayes' rule 

- Law of large numbers 
  - Central limit theorem

- Probability distributions 
  - continuous 
  - discrete
  

## Additional resources 

### Cheatsheets 

- [Probability cheatsheet](figures/probability.pdf)

### Data camp courses 

- [Foundations of probability in R](https://www.datacamp.com/courses/foundations-of-probability-in-r)

### Books and chapters

- [Probability and Statistics with examples using R](http://www.isibang.ac.in/~athreya/psweur/)
- [Learning statistics with R: Chapter 9 Introduction to probability](https://learningstatisticswithr-bookdown.netlify.com/probability.html#probstats)

### Tutorials 

### Misc 

- [Statistics 110: Probability; course at Harvard](https://projects.iq.harvard.edu/stat110)  
  
## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r}
sessionInfo()
```
