---
title: "08_Simulation2"
author: "Tobias Gerstenberg"
date: "1/25/2019"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

> Central limit theorem was a little bit confusing......

> Good explanation. I haven't really understood the CLT before

> I think you spent too long on CLT which isn't intuitively difficult and would like more time on the harder topics near the end

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-8.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```


## What to do next time

- this lecture needs some work 
- drop the gender height example -- it's more confusing than helpful 
- better introduce the three key distributions: sample distribution, sampling distribution, population distribution
- visualization and examples for CLT work well otherwise
- maybe: shift the balance (less time on CLT and more time on p-values and confidence intervals)