---
title: "Class 4"
author: "Tobias Gerstenberg"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(comment = "")
```

# Data wrangling 1: Exercise solutions 

## Load packages and data set 

Let's first load the packages that we need for this chapter. 

```{r load_packages, message=FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
```

And let's load the data set into the environment

```{r data}
df.starwars = starwars
```

## Practice 1 

Let's practice the pipe operator. 

```{r practice1-1}
# here are some numbers
x = seq(from = 1, to = 5, by = 1)

# taking the log the standard way
log(x)

# now take the log the pipe way (write your code underneath)
x %>% log()
```

```{r practice1-2}
# some more numbers
x = seq(from = 10, to = 5, by = -1)

# the standard way
mean(round(sqrt(x), digits = 2))

# the pipe way (write your code underneath)
x %>% 
  sqrt() %>% 
  round(digits = 2) %>% 
  mean()
```

## Practice 2

Create a data frame that: 
- only has the species `Human` and `Droid` 
- with the following data columns (in this order): name, species, birth_year, homeworld
- is arranged according to birth year (with the lowest entry at the top of the data frame)
- and has the `name` column renamed to `person`

```{r practice2}
df.starwars %>% 
  filter(species %in% c("Human", "Droid")) %>% 
  select(name, species, birth_year, homeworld) %>% 
  arrange(birth_year) %>% 
  rename(person = name)
```

## Practice 3 

Compute the body mass index for `masculine` characters who are `Human`.

- select only the columns you need 
- filter out only the rows you need 
- make the new variable with the body mass index 
- arrange the data frame starting with the highest body mass index 

```{r practice3}
df.starwars %>% 
  select(species, gender, height, mass) %>% 
  filter(gender == "masculine", species == "Human") %>% 
  mutate(bmi = mass / (height/100)^2) %>% 
  arrange(desc(bmi))
```
