---
title: "04_Data_wrangling1"
author: "Tobias Gerstenberg"
date: "1/14/2019"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

> Thanks for the class! I know that you had a lot to get through, but I would appreciate more exercises (like the ones we had last week) so the class is a bit more interactive.

> Everything is good and pace is just right. The only thing that I am concern is getting the practice answers from the lectures and sessions. Will it be possible for the lecturers and tutors to __post the practice answers online__, so that people who may have accidentially missed the answers can check it after classes? Thanks!

> I think a lot of today's class could have been abbreviated by having folks read through the file before class. This would have left more time for questions/exercises.

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-4.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read.csv("how-was-the-pace-of-4.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
ggsave("pace.pdf", width = 8, height = 6)
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-4.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
ggsave("overall.pdf", width = 8, height = 6)
```

## What to do next time

- include some more practice problems (fun stuff to play around with for more advanced users so they don't get bored)
- post answers to questions online 
- encourage students to take a look at the materials before class 
