---
title: "04_Data_wrangling1"
author: "Tobias Gerstenberg"
date: "1/13/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(theme_classic() +
            theme(text = element_text(size = 20),
                  panel.grid.major.y = element_line(color = "gray90")))
```

## Student feedback 

> The lecture part was a bit long and there was not enough practice time for the problems.

> I think 5 minutes was too much time for the first two practice exercises.

> I thought the class went well. The explanations I valued most were the reasons why things did not work.

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-4.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

<!-- ### PollEverywhere results  -->

<!-- ```{r echo=FALSE, message=FALSE, results='hide'} -->
<!-- df.pace = read.csv("how-was-the-pace-of-4.csv", header = F) %>% -->
<!--   filter(row_number() > which(.$V2 == "Created At")) %>%  -->
<!--   set_names(c("response", "date")) %>%  -->
<!--   mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"), -->
<!--                            labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>%  -->
<!--   count(response) %>%  -->
<!--   complete(response, fill = list(n = 0)) -->

<!-- ggplot(data = df.pace, -->
<!--        mapping = aes(x = response, y = n)) +  -->
<!--   geom_bar(stat = "identity", fill = "lightblue", color = "black") + -->
<!--   labs(y = "number of students", x = "", title = "How was the pace of today's class?") -->
<!-- ggsave("pace.pdf", width = 8, height = 6) -->
<!-- ``` -->

<!-- ```{r echo=FALSE, message=FALSE, results='hide'} -->
<!-- df.overall = read_csv("how-happy-were-you-with-4.csv") %>%  -->
<!--   set_names(c("response", "date")) %>%  -->
<!--   filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions  -->
<!--   mutate(response = as.numeric(response), -->
<!--          response = factor(response, levels = 1:5, -->
<!--                            labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>%  -->
<!--   count(response) %>%  -->
<!--   complete(response, fill = list(n = 0)) -->

<!-- ggplot(data = df.overall, -->
<!--        mapping = aes(x = response, y = n)) +  -->
<!--   geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) + -->
<!--   scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) + -->
<!--   labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") + -->
<!--   theme(title = element_text(size = 18)) -->
<!-- ggsave("overall.pdf", width = 8, height = 6) -->
<!-- ``` -->

## What to do next time

- expand problem 1 (or give less time for it)
- maybe: go through more examples where stuff fails, and then explain why 
