---
title: "Class 23"
author: "Tobias Gerstenberg"
date: ""
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=22"]
bibliography: references.bib
biblio-style: apalike
---

# Bayesian data analysis 3

## Learning goals 

- Evidence for null results. 
- Only positive predictors. 
- Dealing with unequal variance. 
- Modeling slider data: Zero-one inflated beta binomial model. 
- Modeling Likert scale data: Ordinal logistic regression. 

## Load packages and set plotting theme  

```{r, message=FALSE}
library("knitr")       # for knitting RMarkdown 
library("kableExtra")  # for nice RMarkdown tables
library("janitor")     # for cleaning column names
library("modelr")      # for doing modeling stuff
library("tidybayes")   # tidying up results from Bayesian models
library("brms")        # Bayesian regression models with Stan
library("rstanarm")    # for Bayesian models
library("patchwork")   # for making figure panels
library("ggrepel")     # for labels in ggplots
library("broom.mixed") # for tidying things up
library("gganimate")   # for animations
library("GGally")      # for pairs plot
library("patchwork")   # for figure panels
library("parameters")  # for extracting parameters
library("bayesplot")   # for visualization of Bayesian model fits 
library("ggeffects")   # for showing marginal/conditional effects
library("scales")      # for percent y-axis
library("tidyverse")   # for wrangling, plotting, etc. 
```

```{r}
theme_set(theme_classic() + #set the theme 
            theme(text = element_text(size = 20))) #set the default text size

# set rstan options
rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

opts_chunk$set(comment = "",
               fig.show = "hold")
```

## Evidence for the null hypothesis 

See [this tutorial](https://vuorre.netlify.com/post/2017/03/21/bayes-factors-with-brms/) and this paper [@wagenmakers2010bayesiana] for more information. 

### Bayes factor

#### Fit the model 

- Define a binomial model
- Give a uniform prior `beta(1, 1)`
- Get samples from the prior

```{r}
df.null = tibble(s = 6, k = 10)

fit.brm_bayes = brm(s | trials(k) ~ 0 + Intercept, 
               family = binomial(link = "identity"),
               prior = set_prior("beta(1, 1)", class = "b", lb = 0, ub = 1),
               data = df.null,
               sample_prior = TRUE,
               cores = 4,
               file = "cache/brm_bayes")
```

#### Visualize the results 

Visualize the prior and posterior samples: 

```{r}
fit.brm_bayes %>% 
  posterior_samples(pars = "b") %>% 
  pivot_longer(cols = everything()) %>% 
  ggplot(mapping = aes(x = value,
                       fill = name)) + 
  geom_density(alpha = 0.5) + 
  scale_fill_brewer(palette = "Set1")
```


We test the H0: $\theta = 0.5$ versus the H1: $\theta \neq 0.5$ using the Savage-Dickey Method, according to which we can compute the Bayes factor like so:  

$BF_{01} = \frac{p(D|H_0)}{p(D|H_1)} = \frac{p(\theta = 0.5|D, H_1)}{p(\theta = 0.5|H_1)}$

```{r}
fit.brm_bayes %>% 
  hypothesis(hypothesis = "Intercept = 0.5")
```

The result shows that the evidence ratio is in favor of the H0 with $BF_{01} = 2.22$. This means that H0 is 2.2 more likely than H1 given the data. 

### LOO 

Another way to test different models is to compare them via approximate leave-one-out cross-validation. 

```{r}
set.seed(1)
df.loo = tibble(x = rnorm(n = 50),
                y = rnorm(n = 50))

# visualize 
ggplot(data = df.loo,
       mapping = aes(x = x, 
                     y = y)) + 
  geom_point()

# fit the frequentist model 
fit.lm_loo = lm(formula = y ~ 1 + x,
                data = df.loo)

fit.lm_loo %>% 
  summary()

# fit and compare bayesian models 
fit.brm_loo1 = brm(formula = y ~ 1,
                   data = df.loo,
                   seed = 1, 
                   file = "cache/brm_loo1")

fit.brm_loo2 = brm(formula = y ~ 1 + x,
                   data = df.loo,
                   seed = 1, 
                   file = "cache/brm_loo2")

fit.brm_loo1 = add_criterion(fit.brm_loo1,
                             criterion = "loo",
                             file = "cache/brm_loo1")

fit.brm_loo2 = add_criterion(fit.brm_loo2,
                             criterion = "loo",
                             file = "cache/brm_loo2")

loo_compare(fit.brm_loo1, fit.brm_loo2)
model_weights(fit.brm_loo1, fit.brm_loo2)
```


## Dealing with heteroscedasticity 

Let's generate some fake developmental data where the variance in the data is greatest for young children, smaller for older children, and even smaller for adults:  

```{r}
# make example reproducible 
set.seed(1)

df.variance = tibble(group = rep(c("3yo", "5yo", "adults"), each = 20),
                     response = rnorm(n = 60,
                                      mean = rep(c(0, 5, 8), each = 20),
                                      sd = rep(c(3, 1.5, 0.3), each = 20)))

```

### Visualize the data 

```{r}
df.variance %>%
  ggplot(aes(x = group, y = response)) +
  geom_jitter(height = 0,
              width = 0.1,
              alpha = 0.7)
```

### Frequentist analysis 

#### Fit the model

```{r}
fit.lm_variance = lm(formula = response ~ 1 + group,
                     data = df.variance)

fit.lm_variance %>% 
  summary()

fit.lm_variance %>% 
  glance()
```

#### Visualize the model predictions 

```{r}
set.seed(1)

fit.lm_variance %>% 
  simulate() %>% 
  bind_cols(df.variance) %>% 
  ggplot(aes(x = group, y = sim_1)) +
  geom_jitter(height = 0,
              width = 0.1,
              alpha = 0.7)
```

Notice how the model predicts that the variance is equal for each group.

### Bayesian analysis 

While frequentist models (such as a linear regression) assume equality of variance, Bayesian models afford us with the flexibility of inferring both the parameter estimates of the groups (i.e. the means and differences between the means), as well as the variances. 

#### Fit the model 

We define a multivariate model which tries to fit both the `response` as well as the variance `sigma`: 

```{r}
fit.brm_variance = brm(formula = bf(response ~ group,
                                    sigma ~ group),
                       data = df.variance,
                       file = "cache/brm_variance",
                       seed = 1)

summary(fit.brm_variance)
```

Notice that sigma is on the log scale. To get the standard deviations, we have to exponentiate the predictors, like so:  

```{r}
fit.brm_variance %>% 
  tidy(parameters = "^b_") %>% 
  filter(str_detect(term, "sigma")) %>% 
  select(term, estimate) %>% 
  mutate(term = str_remove(term, "b_sigma_")) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  clean_names() %>% 
  mutate(across(-intercept, ~ exp(. + intercept))) %>% 
  mutate(intercept = exp(intercept))
```

#### Visualize the model predictions 

```{r}
df.variance %>%
  expand(group) %>% 
  add_fitted_draws(fit.brm_variance, dpar = TRUE) %>%
  select(group, .row, .draw, posterior = .value, mu, sigma) %>% 
  pivot_longer(cols = c(mu, sigma),
               names_to = "index",
               values_to = "value") %>% 
  ggplot(aes(x = value, y = group)) +
  geom_halfeyeh() +
  geom_vline(xintercept = 0, linetype = "dashed") +
  facet_grid(cols = vars(index))
```

This plot shows what the posterior looks like for both mu (the inferred means), and for sigma (the inferred variances) for the different groups. 

```{r}
set.seed(1)

df.variance %>% 
  add_predicted_draws(model = fit.brm_variance,
                      n = 1) %>% 
  ggplot(aes(x = group, y = .prediction)) +
  geom_jitter(height = 0,
              width = 0.1,
              alpha = 0.7)
```

## Zero-one inflated beta binomial model 

See this [blog post](https://vuorre.netlify.com/post/2019/02/18/analyze-analog-scale-ratings-with-zero-one-inflated-beta-models/#zoib-regression). 

## Ordinal regression 

Check out the following two papers: 

- @liddell2018analyzin
- @burkner2019ordinal

Let's read in some movie ratings: 

```{r, warning=F, message=F}
df.movies = read_csv(file = "data/MoviesData.csv")

df.movies = df.movies %>% 
  pivot_longer(cols = n1:n5,
               names_to = "stars",
               values_to = "rating") %>% 
  mutate(stars = str_remove(stars,"n"),
         stars = as.numeric(stars))

df.movies = df.movies %>% 
  uncount(weights = rating) %>% 
  mutate(id = as.factor(ID)) %>% 
  filter(ID <= 6)
```

### Ordinal regression (assuming equal variance)

#### Fit the model 

```{r}
fit.brm_ordinal = brm(formula = stars ~ 1 + id,
                      family = cumulative(link = "probit"),
                      data = df.movies,
                      file = "cache/brm_ordinal",
                      seed = 1)

summary(fit.brm_ordinal)
```

#### Visualizations 

##### Model parameters 

The model infers the thresholds and the means of the Gaussian distributions in latent space. 

```{r, warning=FALSE, message=FALSE}
df.params = fit.brm_ordinal %>% 
  parameters(centrality = "mean") %>% 
  as_tibble() %>% 
  clean_names() %>% 
  select(term = parameter, estimate = mean)

ggplot(data = tibble(x = c(-3, 3)),
       mapping = aes(x = x)) + 
  stat_function(fun = ~ dnorm(.),
                size = 1,
                color = "black") +
  stat_function(fun = ~ dnorm(., mean = df.params %>% 
                                filter(str_detect(term, "id2")) %>% 
                                pull(estimate)),
                size = 1,
                color = "blue") +
  geom_vline(xintercept = df.params %>% 
               filter(str_detect(term, "Intercept")) %>% 
               pull(estimate))
```

##### MCMC inference 

```{r, fig.height=20, fig.width=8}
fit.brm_ordinal %>% 
  plot(N = 9)
```

```{r}
fit.brm_ordinal %>% 
  pp_check(nsamples = 20)
```


##### Model predictions 

```{r}
conditional_effects(fit.brm_ordinal,
                    effects = "id",
                    categorical = T)
```

```{r}
df.model = add_fitted_draws(newdata = expand_grid(id = 1:6),
                           model = fit.brm_ordinal,
                           n = 10)

df.plot = df.movies %>% 
  count(id, stars) %>% 
  group_by(id) %>% 
  mutate(p = n / sum(n)) %>% 
  mutate(stars = as.factor(stars))
  
ggplot(data = df.plot,
       mapping = aes(x = stars,
                     y = p)) +
  geom_col(color = "black",
           fill = "lightblue") +
  geom_point(data = df.model,
             mapping = aes(x = .category,
                           y = .value),
             alpha = 0.3,
             position = position_jitter(width = 0.3)) +
  facet_wrap(~id, ncol = 6) 
```

### Gaussian regression (assuming equal variance)

#### Fit the model 

```{r}
fit.brm_metric = brm(formula = stars ~ 1 + id,
               data = df.movies,
               file = "cache/brm_metric",
               seed = 1)

summary(fit.brm_metric)
```

#### Visualizations 

##### Model predictions 

```{r, message=FALSE}
# get the predictions for each value of the Likert scale 
df.model = fit.brm_metric %>% 
  parameters(centrality = "mean") %>% 
  as_tibble() %>% 
  select(term = Parameter, estimate = Mean) %>% 
  mutate(term = str_remove(term, "b_")) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  clean_names() %>%
  mutate(across(id2:id6, ~ . + intercept)) %>% 
  set_names(str_c("mu_", 1:6)) %>%
  pivot_longer(cols = everything(),
               names_to = c("parameter", "movie"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = parameter, 
              values_from = value) %>% 
  mutate(data = map(.x = mu, .f = ~ tibble(x = 1:5,
                                           y  = dnorm(x, mean = .x)))) %>% 
  select(movie, data) %>% 
  unnest(c(data)) %>% 
  group_by(movie) %>% 
  mutate(y = y/sum(y)) %>% 
  ungroup() %>% 
  rename(id = movie)

# visualize the predictions 
df.plot = df.movies %>% 
  count(id, stars) %>% 
  group_by(id) %>% 
  mutate(p = n / sum(n)) %>% 
  mutate(stars = as.factor(stars))

ggplot(data = df.plot,
       mapping = aes(x = stars,
                     y = p)) +
  geom_col(color = "black",
           fill = "lightblue") +
  geom_point(data = df.model,
            mapping = aes(x = x,
                          y = y)) +
  facet_wrap(~id, ncol = 6) 
```

### Oridnal regression (unequal variance) 

#### Fit the model 

```{r}
fit.brm_ordinal_variance = brm(formula = bf(stars ~ 1 + id) + lf(disc ~ 0 + id, cmc = FALSE),
               family = cumulative(link = "probit"),
               data = df.movies,
               file = "cache/brm_ordinal_variance",
               seed = 1)

summary(fit.brm_ordinal_variance)
```

#### Visualizations 

##### Model parameters 

```{r}
df.params = fit.brm_ordinal_variance %>% 
  tidy(parameters = "^b_") %>% 
  select(term, estimate) %>% 
  mutate(term = str_remove(term, "b_"))

ggplot(data = tibble(x = c(-3, 3)),
       mapping = aes(x = x)) + 
  stat_function(fun = ~ dnorm(.),
                size = 1,
                color = "black") +
  stat_function(fun = ~ dnorm(.,
                              mean = 1,
                              sd = 2),
                size = 1,
                color = "blue") +
  geom_vline(xintercept = df.params %>% 
               filter(str_detect(term, "Intercept")) %>% 
               pull(estimate))
```

##### Model predictions 

```{r}
df.model = add_fitted_draws(newdata = expand_grid(id = 1:6),
                           model = fit.brm_ordinal_variance,
                           n = 10)

df.plot = df.movies %>% 
  count(id, stars) %>% 
  group_by(id) %>% 
  mutate(p = n / sum(n)) %>% 
  mutate(stars = as.factor(stars))
  
ggplot(data = df.plot,
       mapping = aes(x = stars,
                     y = p)) +
  geom_col(color = "black",
           fill = "lightblue") +
  geom_point(data = df.model,
             mapping = aes(x = .category,
                           y = .value),
             alpha = 0.3,
             position = position_jitter(width = 0.3)) +
  facet_wrap(~id, ncol = 6) 
```

### Gaussian regression (unequal variance)

#### Fit the model 

```{r}
fit.brm_metric_variance = brm(formula = bf(stars ~ 1 + id,
                            sigma ~ 1 + id),
               data = df.movies,
               file = "cache/brm_metric_variance",
               seed = 1)

summary(fit.brm_metric_variance)
```

#### Visualizations 

##### Model predictions 

```{r}
df.model = fit.brm_metric_variance %>% 
  tidy(parameters = "^b_") %>% 
  select(term, estimate) %>% 
  mutate(term = str_remove(term, "b_")) %>% 
  pivot_wider(names_from = term,
              values_from = estimate) %>% 
  clean_names() %>%
  mutate(across(c(id2:id6), ~ . + intercept)) %>% 
  mutate(across(contains("sigma"), ~ 1/exp(.))) %>% 
  mutate(across(c(sigma_id2:sigma_id5), ~ . + sigma_intercept)) %>% 
  set_names(c("mu_1", "sigma_1", str_c("mu_", 2:6), str_c("sigma_", 2:6))) %>% 
  pivot_longer(cols = everything(),
               names_to = c("parameter", "movie"),
               names_sep = "_",
               values_to = "value") %>% 
  pivot_wider(names_from = parameter, 
              values_from = value) %>% 
  mutate(data = map2(.x = mu, .y = sigma, .f = ~ tibble(x = 1:5,
                                                        y  = dnorm(x,
                                                       mean = .x,
                                                       sd = .y)))) %>% 
  select(movie, data) %>% 
  unnest(c(data)) %>% 
  group_by(movie) %>% 
  mutate(y = y/sum(y)) %>% 
  ungroup() %>% 
  rename(id = movie)

df.plot = df.movies %>% 
  count(id, stars) %>% 
  group_by(id) %>% 
  mutate(p = n / sum(n)) %>% 
  mutate(stars = as.factor(stars))

ggplot(data = df.plot,
       mapping = aes(x = stars,
                     y = p)) +
  geom_col(color = "black",
           fill = "lightblue") +
  geom_point(data = df.model,
            mapping = aes(x = x,
                          y = y)) +
  facet_wrap(~id, ncol = 6) 
```

### Model comparison 

```{r}
# ordinal regression with equal variance 
fit.brm_ordinal = add_criterion(fit.brm_ordinal,
                                criterion = "loo",
                                file = "cache/brm_ordinal")

# Gaussian regression with equal variance
fit.brm_ordinal_variance = add_criterion(fit.brm_ordinal_variance,
                                         criterion = "loo",
                                         file = "cache/brm_ordinal_variance")

loo_compare(fit.brm_ordinal, fit.brm_ordinal_variance)
```

## Additional resources 

- [Tutorial on visualizing brms posteriors with tidybayes](https://mjskay.github.io/tidybayes/articles/tidy-brms.html)
- [Hypothetical outcome plots](https://mucollective.northwestern.edu/files/2018-HOPsTrends-InfoVis.pdf)
- [Visual MCMC diagnostics](https://cran.r-project.org/web/packages/bayesplot/vignettes/visual-mcmc-diagnostics.html#general-mcmc-diagnostics)
- [Visualiztion of different MCMC algorithms](https://chi-feng.github.io/mcmc-demo/)

## Session info 

Information about this R session including which version of R was used, and what packages were loaded.

```{r}
sessionInfo()
```

## References {-}