---
title: "Class 20"
author: "Tobias Gerstenberg"
date: "February 26th, 2020"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=19"]
bibliography: [packages.bib]
nocite: '@*'
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(comment = "#>",
                      fig.show = "hold")
```

# Generalized linear model 

```{r glm-01, include=FALSE, eval=FALSE}
# new packages
install.packages(c("titanic", "effects", "nlme"))
```

## Learning goals 

- Logistic regression. 
- Logit transform. 
- Fitting a logistic regression in R.
- Visualizing and interpreting model predictions. 
- Simulating data from a logistic regression. 
- Assessing model fit. 
- Testing hypotheses. 
- Reporting results.
- Mixed effects logistic regression.

## Load packages and set plotting theme  

```{r glm-02, message=FALSE}
library("knitr")      # for knitting RMarkdown 
library("titanic")    # titanic dataset
library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")      # for tidying up linear models 
library("lme4")       # for linear mixed effects models
library("boot")       # for bootstrapping (also has an inverse logit function)
library("ggeffects")  # for showing effects in linear, generalized linear, and other models
library("tidyverse")  # for wrangling, plotting, etc. 

# include references for used packages
knitr::write_bib(.packages(), "packages.bib") 
```

```{r glm-03}
theme_set(theme_classic() + #set the theme 
    theme(text = element_text(size = 20))) #set the default text size
```

## Load data set 

```{r glm-04}
df.titanic = titanic_train %>% 
  clean_names() %>% 
  mutate(sex = as.factor(sex))
```

Let's take a quick look at the data: 

```{r glm-05}
df.titanic %>% 
  glimpse()
```

```{r glm-06}
# Table of the first 10 entries
df.titanic %>% 
  head(10) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
```

## Logistic regression 

Let's see if we can predict whether or not a passenger survived based on the price of their ticket. 

Let's run a simple regression first: 

```{r glm-07}
# fit a linear model 
fit.lm = lm(formula = survived ~ 1 + fare,
            data = df.titanic)

# summarize the results
fit.lm %>% summary()
```

Look's like `fare` is a significant predictor of whether or not a person survived. Let's visualize the model's predictions:

```{r glm-08}
ggplot(data = df.titanic,
       mapping = aes(x = fare,
                     y = survived)) + 
  geom_smooth(method = "lm") + 
  geom_point() +
  labs(y = "survived")
```

This doesn't look good! The model predicts intermediate values of `survived` (which doesn't make sense given that a person either survived or didn't survive). Furthermore, the model predicts values greater than 1 for fares greather than ~ 300.  

Let's run a logistic regression instead. 

```{r glm-09}
# fit a logistic regression 
fit.glm = glm(formula = survived ~ 1 + fare,
              family = "binomial",
              data = df.titanic)

fit.glm %>% 
  summary()
```

And let's visualize the predictions of the logistic regression: 

```{r glm-10}
ggplot(data = df.titanic,
       mapping = aes(x = fare,
                     y = survived)) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial")) + 
  geom_point() +
  labs(y = "p(survived)")
```

Much better! Note that we've changed the interpretation of our dependent variable. We are now predicting the _probability that a person survived_ based on their fare. The model now only predicts values between 0 and 1. To achieve this, we apply a logit transform to the outcome variable like so: 

$$
\ln(\frac{\pi_i}{1-\pi_i}) = b_0 + b_1 \cdot X_i + e_i
$$
where $\pi_i$ is the probability of passenger $i$ having survived. Importantly, this affects our interpretation of the model parameters. They are now defined in log-odds, and can apply an inverse logit transformation to turn this back into a probability: 

With

$$
\pi = P(Y = 1)
$$
and the logit transformation 

$$
\ln(\frac{\pi}{1-\pi}) = V,
$$
where $V$ is just a placeholder for our linear model formula, we can go back to $\pi$ through the inverse logit transformation like so: 

$$
\pi = \frac{e^V}{1 + e^V}
$$
In R, we can use `log(x)` to calculate the natural logarithm $\ln(x)$, and `exp(x)` to calculate `e^x`. 

### Interpreting the parameters 

```{r glm-11}
fit.glm %>% 
  summary()
```

The estimate for the intercept and fare are in log-odds. We apply the inverse logit transformation to turn these into probabilities: 

```{r glm-12}
fit.glm$coefficients[1] %>% 
  inv.logit()
```

Here, we see that the intercept is $p = 0.28$. That is, the predicted chance of survival for someone who didn't pay any fare at all is 28% according to the model. Interpreting the slope is a little more tricky. Let's look at a situation first where we have a binary predictor. 

#### Binary predictor

Let's see whether the probability of survival differed between male and female passengers. 

```{r glm-13}
fit.glm2 = glm(formula = survived ~ sex,
               family = "binomial",
               data = df.titanic)

fit.glm2 %>% 
  summary()
```

It looks like it did! Let's visualize: 

```{r glm-14}
df.titanic %>% 
  mutate(survived = factor(survived, labels = c("died", "survived"))) %>% 
  ggplot(data = .,
         mapping = aes(x = sex,
                       fill = survived)) +
  geom_bar(position = "fill",
           color = "black") +
  scale_fill_brewer(palette = "Set1")  +
  labs(x = "", fill = "", y = "probability")
  
```

And let's interpret the parameters by applying the inverse logit transform. To get the prediction for female passengers we get 

$$
\widehat{\ln(\frac{\pi_i}{1-\pi_i})} = b_0 + b_1 \cdot \text{sex}_i = b_0 + b_1 \cdot 0 = b_0
$$

since we dummy coded the predictor and female is our reference category. To get the predicted probability of surival for women we do the logit transform: 

$$
\pi = \frac{e^{b_0}}{1 + e^{b_0}}
$$
The predicted probability is: 

```{r glm-15}
fit.glm2$coefficients[1] %>% inv.logit()
```

To get the prediction for male passengers we have: 

$$
\widehat{\ln(\frac{\pi_i}{1-\pi_i})} = b_0 + b_1 \cdot \text{sex}_i = b_0 + b_1 \cdot 1 = b_0 + b_1
$$
Applying the logit transform like so

$$
\pi = \frac{e^{b_0 + b_1}}{1 + e^{b_0 + b_1}}
$$

The predicted probability of male passengers surviving is: 

```{r glm-16}
sum(fit.glm2$coefficients) %>% inv.logit()
```

Here is the same information in a table: 

```{r glm-17}
df.titanic %>% 
  count(sex, survived) %>% 
  mutate(p = n/sum(n)) %>% 
  group_by(sex) %>% 
  mutate(`p(survived|sex)` = p/sum(p)) %>% 
  head(10) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
  
```

#### Continuous predictor

To interpret the predictions when a continuous predictor is invovled, it's easiest to consider a few concrete cases. Here, I use the `augment()` function from the "broom" package to get the model's predictions for some values of interest: 

```{r glm-18}
fit.glm %>% 
  augment(newdata = tibble(fare = c(0, 10, 50, 100, 500))) %>% 
  clean_names() %>% 
  select(fare, prediction = fitted) %>% 
  mutate(`p(survival)` = prediction %>% inv.logit()) %>% 
  head(10) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
  
```

#### Several predictors 

Let's fit a logistic regression that predicts the probability of survival based both on the passenger's sex and what fare they paid (allowing for an interaction of the two predictors): 

```{r glm-19}
fit.glm = glm(formula = survived ~ 1 + sex + fare,
              family = "binomial",
              data = df.titanic)

fit.glm %>% 
  summary()
```

And let's visualize the result: 

```{r glm-20}
df.titanic %>% 
  mutate(sex = as.factor(sex)) %>% 
  ggplot(data = .,
         mapping = aes(x = fare,
                       y = survived,
                       color = sex)) +
  geom_point(alpha = 0.1, size = 2) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              alpha = 0.2,
              aes(fill = sex)) +
  scale_color_brewer(palette = "Set1")
```

Let's see how large the difference between genders is once we take into account how much each person paid for the fair:

```{r glm-21} 
ggpredict(fit.glm,
          terms = c("sex"))
```


We notice that there is one outlier who was male and paid a $500 fare (or maybe this is a mistake in the data entry?!). Let's remove this outlier and see what happens: 

```{r glm-22}
fit.glm2 = glm(formula = survived ~ 1 + sex + fare,
              family = "binomial",
              data = df.titanic %>% 
                filter(fare < 500))

fit.glm2 %>% 
  summary()

df.titanic %>% 
  filter(fare < 500) %>% 
  mutate(sex = as.factor(sex)) %>% 
  ggplot(data = .,
         mapping = aes(x = fare,
                       y = survived,
                       color = sex)) +
  geom_point(alpha = 0.1, size = 2) + 
  stat_smooth(method = "glm",
              method.args = list(family = "binomial"),
              alpha = 0.2,
              fullrange = T,
              aes(fill = sex)) +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 500))
```

There is still a clear difference between female and male passengers, but the prediction for male passengers has changed a bit. Let's look at a concrete example: 

```{r glm-23}
# with the outlier: 

# predicted probability of survival for a male passenger who paid $200 for their fare 
inv.logit(fit.glm$coefficients[1] + fit.glm$coefficients[2] + fit.glm$coefficients[3] * 200 + fit.glm$coefficients[4] * 200)
  

# without the outlier: 

# predicted probability of survival for a male passenger who paid $200 for their fare 
fit.glm_no_outlier = glm(formula = survived ~ 1 + sex * fare,
                         family = "binomial",
                         data = df.titanic %>% 
                           filter(fare < 500))

inv.logit(fit.glm_no_outlier$coefficients[1] + fit.glm_no_outlier$coefficients[2] + fit.glm_no_outlier$coefficients[3] * 200 + fit.glm_no_outlier$coefficients[4] * 200) 
```

With the oulier removed, the predicted probability of survival for a male passenger who paid $200 decreases from 49% to 47%. 

##### With interactions 

```{r glm-24} 
fit.glm3 = glm(formula = survived ~ 1 + sex * fare,
              family = "binomial",
              data = df.titanic)

df.data = ggpredict(fit.glm3,
                    terms = c("fare [0:500]", "sex"))
  
ggplot(data = df.data,
       mapping = aes(x = x,
                     y = predicted,
                     color = group)) +
  geom_ribbon(mapping = aes(ymin = conf.low,
                            ymax = conf.high,
                            fill = group),
              alpha = 0.2,
              color = NA) +
  geom_line(size = 1) + 
  geom_point(data = df.titanic,
             mapping = aes(x = fare,
                           y = survived,
                           color = sex),
             alpha = 0.2) + 
  scale_color_brewer(palette = "Set1")
```



#### Using the "ggeffects" package 

The "ggeffects" package helps with the interpretation of the results. It applies the inverse logit transform for us, and shows the predictions for a range of cases. 

```{r glm-25}
# show effects 
ggeffect(model = fit.glm,
         terms = "fare [0, 100, 200, 300, 400, 500]")
```

I've used the `[]` construction to specify for what values of the predictor `fare`, I'd like get the predicted values. 

## Simulate a logistic regression

As always, to better understand a statistical modeling procedure, it's helpful to simulate data from the assumed data-generating process, fit the model, and see whether we can reconstruct the parameters.  

```{r glm-26}
# make example reproducible 
set.seed(1)

# set parameters 
sample_size = 1000 
b0 = 0
b1 = 1
# b1 = 8

# generate data 
df.data = tibble(
  x = rnorm(n = sample_size),
  y = b0 + b1 * x,
  p = inv.logit(y)) %>% 
  mutate(response = rbinom(n(), size = 1, p = p))

# fit model 
fit = glm(formula = response ~ 1 + x,
          family = "binomial",
          data = df.data)

# model summary 
fit %>% summary()
```

Nice! The inferred estimates are very close to the parameter values we used to simulate the data. 

Let's visualize the result: 

```{r glm-27}
ggplot(data = df.data,
       mapping = aes(x = x,
                     y = response)) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial")) + 
  geom_point(alpha = 0.1) +
  labs(y = "p(response)")
```

#### Calculate the model's likelihood 

To calculate the likelihood of the data for a given logistic model, we look at the actual response, and the probability of the predicted response, and then determine the likelihood of the observation assuming a bernoulli process. To get the overall likelihood of the data, we then multiply the likelihood of each data point (or take the logs first and then the sum to get the log-likelihood). 

This table illustrate the steps involved: 

```{r glm-28}
fit %>% 
  augment() %>% 
  clean_names() %>% 
  mutate(p = inv.logit(fitted)) %>% 
  select(response, p) %>% 
  mutate(p_response = ifelse(response == 1, p, 1-p),
         log_p = log(p_response)) %>% 
  rename(`p(Y = 1)` = p, `p(Y = response)` = p_response,
         `log(p(Y = response))` = log_p) %>% 
  head(10) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
```

Let's calculate the log-likelihood by hand:

```{r glm-29}
fit %>% 
  augment() %>% 
  clean_names() %>% 
  mutate(p = inv.logit(fitted),
         log_likelihood = response * log(p) + (1 - response) * log(1 - p)) %>% 
  summarize(log_likelihood = sum(log_likelihood))

```

And compare it with the model summary

```{r glm-30}
fit %>% 
  glance() %>% 
  select(logLik, AIC, BIC) %>% 
  kable(digits = 2) %>% 
  kable_styling(bootstrap_options = "striped",
                full_width = F)
  
```

We're getting the same result -- neat! Now we know how the likelihood of the data is calculated for a logistic regression model. 

## Testing hypotheses

To test hypotheses, we use our gold old model comparison approach: 

```{r glm-31}
# fit compact model
fit.compact = glm(formula = survived ~ 1 + fare,
                  family = "binomial",
                  data = df.titanic)

# fit augmented model
fit.augmented = glm(formula = survived ~ 1 + sex + fare,
                    family = "binomial",
                    data = df.titanic)

# likelihood ratio test
anova(fit.compact, fit.augmented, test = "LRT")

```

Note that in order to get a p-value out of this, we need to specify what statistical test we'd like to run. In this case, we use the likelihood ratio test ("LRT"). 

## Logistic mixed effects model 

Just like we can build linear mixed effects models using `lmer()` instead of `lm()`, we can also build a logistic mixed effects regression using `glmer()` instead of `glm()`. 

Let's read in some data: 

```{r glm-32}
# load bdf data set from nlme package
data(bdf, package = "nlme")

df.language = bdf %>% 
  clean_names() %>% 
  filter(repeatgr != 2) %>% 
  mutate(repeatgr = repeatgr %>% as.character() %>% as.numeric())

rm(bdf)
```

Fit the model, and print out the results: 

```{r glm-33}
fit =  glmer(repeatgr ~ 1 + ses + minority + (1 | school_nr),
             data = df.language,
             family = "binomial")

fit %>%
  summary()
```

To visualize the results, we can use the `ggeffects` package. 

```{r glm-34} 
ggpredict(model = fit,
          terms = c("ses [all]", "minority")) %>% 
  plot()
```


## Logit transform 

Here is a visualization of what the odds and log odds transformation look like.

```{r glm-35} 
# odds 
ggplot(data = tibble(x = seq(0, 1, 0.1)),
       mapping = aes(x = x)) + 
  stat_function(fun = ~ ./(1 - .),
                size = 1)

# logg odds 
ggplot(data = tibble(x = seq(0, 1, 0.1)),
       mapping = aes(x = x)) + 
  stat_function(fun = ~ log(./(1 - .)),
                size = 1)

```


## Additional information 

### Datacamp 

- [Multiple and logistic regression](https://www.datacamp.com/courses/multiple-and-logistic-regression)
- [Generalized linear models in R](https://www.datacamp.com/courses/generalized-linear-models-in-r)
- [Categorical data in the tidyverse](https://www.datacamp.com/courses/categorical-data-in-the-tidyverse)

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r glm-36}
sessionInfo()
```

## References