---
title: "05_Data_wrangling2"
author: "Tobias Gerstenberg"
date: "1/15/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(theme_classic() + 
            theme(
              text = element_text(size = 20),
              panel.grid.major.y = element_line(color = "gray90")))
```

## Student feedback 

> For merging data frames, just to be clear, you didn't mention the function "merge" because you think it's not a good function (should I learn something else?) or you just didn't have time today?

> Today's material was extremely helpful and enlightening! Overall good pace as well. Where can we find solutions to the in-class practice problems?

> I appreciate the applied lecture and like all of the practice we had. Although, I would like additional practice in these kinds of lectures because it moved very quickly for me!

> I thought the explanation for some of the practice problems was a little too fast. It is working under the assumption that everyone got it correctly and was just checking their answer as opposed to helping those who may still be a little stuck or missed something figure out a correct solution.

> The explanations of the practice examples went WAY too quickly for me (and for many of the people around me). Especially for those of us who are stuck, the explanation went too quickly for me to absorb the mistakes that I made and how to properly correct them.

> Very helpful & necessary functions - saves life

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-5.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read.csv("how-was-the-pace-of-5.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-5.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
```

## What to do next time

- spend more time on going through the solutions of the practice problems 
- send out an earlier email to students to have them play around with data camp courses to get up to speed 
