---
title: "06_Probability"
author: "Tobias Gerstenberg"
date: "1/17/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

No feedback collected this time 

## Homework 1

```{r echo=FALSE, fig.height=4, fig.width=8, message=FALSE, warning=FALSE, results='hide'}
df.hw1_hours = read.csv("how-many-hours-did-it-1.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response,
                           levels = c(1:14, "15\nor more"))) %>% 
  count(response) %>% 
  complete(response)

ggplot(data = df.hw1_hours,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity",
           fill = "lightblue",
           color = "black") +
  scale_y_continuous(breaks = seq(0, 50, 2),
                     expand = c(0, 0)) +
  labs(y = "number of students", x = "", title = "How many hours did you spend on homework 1?")
```

## What to do next time

- class went well this time 
- added exercise on the secretary problem helped to foster the intuition behind the naive definition of probability (and for using programming / simulation as a way of answering the problem)
- highlighted the importance of getting the intuition behind the patterns of inference that happen in Bayes nets 
- made the point that regression assumes no structure between the predictors (but they are often causally related in the real world)
- maybe: make the point clearer how RCTs are a form of intervention in a CBN 
- maybe: use this shinyapp for illustration in class: https://research-methods-dojo.shinyapps.io/causality_and_regression/