---
title: "03_Visualization2"
author: "Tobias Gerstenberg"
date: "1/10/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

> I wish we could spend a little more time on the content and less time on reviewing feedback. I think it's *great* that you are so interested in soliciting our feedback, but I also want to be sure we can cover the most complicated material in class. Maybe a digest email at the end of the week? Or allow people with specific questions to include their name so they can be contacted directly?

> I'd love to revisit / go into greater depth on stat_summary – I don't have a great sense of the range of ways it can be used. I also would have enjoyed one more practice plot in class!

> I liked that we were exposed to a lot of different types of graphs and other useful things (such as the keyboard shortcuts). I would have liked more time to practice those things together. However, I think we'll get a change to do that in our homeworks and in sections.

> I loved how we got to see the breadth of plots that are possible (bar charts, pie charts, scatterplots, violin plots, etc.). It was just a lot to process, especially for those of us that haven't done stats in a while. If possible, I would have loved to discuss the pros and cons of each type of visualization in a little more detail.

> I learned a ton of useful skills in today's class! No specific suggestions for improvement -- the pace, class material, etc., was all great!


```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-3.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read.csv("how-was-the-pace-of-3.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-3.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
```

## What to do next time

- add another practice plot 
- maybe: spend a little bit more time discussing the pros and cons of different plots 
- didn't have time to show some of the resources at the end 

