---
title: "03_visualization2"
author: "Tobias Gerstenberg"
output: 
  html_document:
    theme: cosmo
    highlight: tango
---

```{r include=FALSE}
library("knitr")
library("DT")
library("janitor")
library("tidyverse")

opts_chunk$set(echo = TRUE)

theme_set(theme_classic() + 
            theme(text = element_text(size = 20),
                  panel.grid.major.y = element_line(color = "gray90")))
```

# Pace

```{r, echo=FALSE}
df.pace = read_csv("data/How was the pace of today's class-.csv",
                   skip = 1,
                   show_col_types = FALSE) %>% 
  clean_names() %>%  
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response,
        .drop = F)

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

# Happiness

```{r, echo=FALSE}
df.happy = read_csv("data/How happy were you with today's class overall-.csv",
                    skip = 1,
                    show_col_types = FALSE) %>% 
  clean_names() %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response,
        .drop = F)

ggplot(data = df.happy,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  scale_y_continuous(breaks = seq(0, 40, 2)) + 
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
```

# What did you like? What can be improved? 

> I appreciate how the instruction is tailored to the skill and experience level of the class, and I think Tobi does a nice job of transitioning from more advanced to more basic topics as needed. It would be helpful to have perhaps an extra minute to work on the practice plots during class.

> Sometimes you finish giving the answer for an activity but does not allow for time so we can fix our own. Would be helpful to just have a few seconds before we move on. Thanks :)


```{r echo=FALSE, message=FALSE}
read_csv("data/What did you like about today's class- What could be improved next time-.csv",
         skip = 1,
         show_col_types = FALSE) %>% 
  datatable()
```

# What to do next time

- leave time for people to fix their code after I showed how to do it and/or remind people that they'll have access to the solutions later 


