---
title: "Class 5"
author: "Tobias Gerstenberg"
date: "January 22nd, 2021"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=4"]
---

# Data wrangling 2: Exercise solutions 

## Load packages and data set 

Let's first load the packages that we need for this chapter. 

```{r load_packages, message=FALSE}
library("knitr") # for rendering the RMarkdown file
library("tidyverse") # for data wrangling 
```

## Settings 

```{r}
opts_chunk$set(comment = "")

options(dplyr.summarise.inform = F)
```


And let's load the data set into the environment

```{r data}
df.starwars = starwars
```

## Practice 1 

Find out what the average `height` and `mass` (as well as the standard deviation) is from different `species` in different `homeworld`s. Why is the standard deviation `NA` for many groups?  

```{r summarize_practice1}
df.starwars %>% 
  group_by(species, homeworld) %>% 
  summarize(mean_height = mean(height, na.rm = T),
            mean_mass = mean(mass, na.rm = T),
            sd_height = sd(height, na.rm = T),
            sd_mass = sd(mass, na.rm = T),
            n = n()) %>% 
  ungroup()
```

The standard deviation is `NA` for many groups since the group size is 1. 

Who is the tallest member of each species? What eye color do they have? The `top_n()` function or the `row_number()` function (in combination with `filter()`) will be useful here. 

```{r summarize_practice2}
df.starwars %>% 
  arrange(species, desc(height)) %>% 
  group_by(species) %>%
  # top_n(1, wt = height) %>% # you can either use this, or filter() in the next row
  filter(row_number() == 1) %>%
  ungroup() %>% 
  select(species, name, eye_color)
```

## Practice 2 

Load this data frame first.

```{r reshaping-practice1}
df.practice2 = tibble(
  participant = 1:10,
  initial = c("AR", "FA", "IR", "NC", "ER", "PI", "DH", "CN", "WT", "JD"), 
  judgment_1 = c(12, 13, 1, 14, 5, 6, 12, 41, 100, 33),
  judgment_2 = c(2, 20, 10, 89, 94, 27, 29, 19, 57, 74),
  judgment_3 = c(2, 20, 10, 89, 94, 27, 29, 19, 57, 74)
)
```

- Make the `df.practice2` data framey tidy (by turning into a long format).
- Compute the z-score of each participants' judgments (using the `scale()` function).
- Calculate the mean and standard deviation of each participants' z-scored judgments. 
- Notice anything interesting? Think about what [z-scoring](https://www.statisticshowto.datasciencecentral.com/probability-and-statistics/z-score/) does ... 

```{r reshaping-practice2}
df.practice2 %>% 
  pivot_longer(cols = -c(participant, initial),
               names_to = "index",
               values_to = "judgment") %>% 
  group_by(participant) %>% 
  mutate(judgment = scale(judgment)) %>% 
  summarize(mean = mean(judgment),
            sd = sd(judgment))
```

## Practice 3

Load these two data frames first: 

```{r joining-practice1}
set.seed(1)

df.judgments = tibble(participant = rep(1:3, each = 5),
                      stimulus = rep(c("red", "green", "blue"), 5),
                      judgment = sample(0:100,
                                        size = 15,
                                        replace = T))

df.information = tibble(number = seq(from = 0,
                                     to = 100,
                                     length.out = 5),
                        color = c("red", "green", "blue", "black", "white"))
```

Create a new data frame called `df.join` that combines the information from both `df.judgments` and `df.information`. Note that column with the colors is called `stimulus` in `df.judgments` and `color` in `df.information`. At the end, you want a data frame that contains the following columns: `participant`, `stimulus`, `number`, and `judgment`. 

```{r joining-practice2}
df.join = df.judgments %>% 
  left_join(df.information,
            by = c("stimulus" = "color")) %>% 
  select(-judgment, everything(), judgment)
df.join
```


