---
title: "Class 19"
author: "Tobias Gerstenberg"
date: "February 25th, 2019"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=18"]
bibliography: [packages.bib]
nocite: '@*'
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(
  comment = "",
  results = "hold",
  fig.show = "hold")
```

# Linear mixed effects models 3

```{r install-packages, include=FALSE, eval=FALSE}
# install.packages(c("lme4", "lmerTest", "pbkrtest"))
```

## Load packages and set plotting theme  

```{r load-packages, message=FALSE}
library("knitr")      # for knitting RMarkdown 
library("kableExtra") # for making nice tables
library("janitor")    # for cleaning column names
library("broom")    # for tidying up linear models 
library("patchwork")    # for making figure panels
library("lme4")    # for linear mixed effects models
library("modelr")    # for bootstrapping
library("boot")    # also for bootstrapping
library("tidyverse")  # for wrangling, plotting, etc. 

# include references for used packages
knitr::write_bib(.packages(), "packages.bib") 
```

```{r set-theme}
theme_set(
  theme_classic() + #set the theme 
    theme(text = element_text(size = 20)) #set the default text size
)
```

## Load data set 

```{r}
# load sleepstudy data set 
df.sleep = sleepstudy %>% 
  as_tibble() %>% 
  clean_names() %>% 
  mutate(subject = as.character(subject)) %>% 
  select(subject, days, reaction)
```

```{r}
# add two fake participants (with missing data)
df.sleep = df.sleep %>% 
  bind_rows(
    tibble(subject = "374",
           days = 0:1,
           reaction = c(286, 288)),
    tibble(subject = "373",
           days = 0,
           reaction = 245)
  )
```

## Things that came up in class 

### One-tailed vs. two-tailed tests

#### t distribution

Some code to draw a t-distribution: 

```{r}
tibble(x = c(-4, 4)) %>% 
  ggplot(data = ., 
         mapping = aes(x = x)) + 
  stat_function(fun = "dt",
                args = list(df = 20),
                size = 1,
                geom = "area",
                fill = "red",
                # xlim = c(qt(0.95, df = 20), qt(0.999, df = 20))) +
                # xlim = c(qt(0.001, df = 20), qt(0.05, df = 20))) +
                xlim = c(qt(0.001, df = 20), qt(0.025, df = 20))) +
  stat_function(fun = "dt",
                args = list(df = 20),
                size = 1,
                geom = "area",
                fill = "red",
                xlim = c(qt(0.975, df = 20), qt(0.999, df = 20))) +
  stat_function(fun = "dt",
                args = list(df = 20),
                size = 1) +
  coord_cartesian(expand = F)
```

#### F distribution

Some code to draw an F-distribution

```{r}
tibble(x = c(0, 5)) %>% 
  ggplot(data = ., 
         mapping = aes(x = x)) +
  stat_function(fun = "df",
                args = list(df1 = 100, df2 = 10),
                size = 1,
                geom = "area",
                fill = "red",
                xlim = c(qf(0.95, df1 = 100, df2 = 10), qf(0.999, df1 = 100, df2 = 10))) +
  stat_function(fun = "df",
                args = list(df1 = 100, df2 = 10),
                size = 1) +
  coord_cartesian(expand = F)
```

### Mixtures of participants 

What if we have groups of participants who differ from each other? Let's generate data for which this is the case.

```{r}
# make example reproducible 
set.seed(1)

sample_size = 20
b0 = 1
b1 = 2
sd_residual = 0.5
sd_participant = 0.5
mean_group1 = 1
mean_group2 = 10

df.mixed = tibble(
  condition = rep(0:1, each = sample_size), 
  participant = rep(1:sample_size, 2)) %>% 
  group_by(participant) %>% 
  mutate(group = sample(1:2, size = 1),
         intercept = ifelse(group == 1,
                            rnorm(n(), mean = mean_group1, sd = sd_participant),
                            rnorm(n(), mean = mean_group2, sd = sd_participant))) %>% 
  group_by(condition) %>% 
  mutate(value = b0 + b1 * condition + intercept + rnorm(n(), sd = sd_residual)) %>% 
  ungroup %>% 
  mutate(condition = as.factor(condition),
         participant = as.factor(participant))
```

#### Ignoring mixture

Let' first fit a model that ignores the fact that there are two different groups of participatns. 

```{r}
# fit model
fit.mixed = lmer(formula = value ~ 1 + condition + (1 | participant),
                data = df.mixed)

fit.mixed %>% summary()
```

Let's look at the model's predictions: 

```{r}
fit.mixed %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```


And let's simulate some data from the fitted model: 

```{r}
# simulated data 
fit.mixed %>%
  simulate() %>%
  bind_cols(df.mixed) %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

As we can see, the simulated data doesn't look like the data that was used to fit the model.  

#### Modeling mixture

Now, let's fit a model that takes the differences between groups into account by adding a fixed effect for `group`.

```{r}
# fit model
fit.grouped = lmer(formula = value ~ 1 + group + condition + (1 | participant),
                data = df.mixed)

fit.grouped %>% summary()
```

Note how the variance of the random intercepts is much smaller now that we've taken the group structure in the data into account. 

Let's visualize the model's predictions:

```{r}
fit.grouped %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```

And simulate some data from the model: 

```{r}
# simulated data 
fit.grouped %>%
  simulate() %>%
  bind_cols(df.mixed) %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

This time, the simulated data looks much more like the data that was used to fit the model. Yay! 

#### Heterogeneity in variance

The example above has shown that we can take overall differences between groups into account by adding a fixed effect. Can we also deal with heterogeneity in variance between groups? For example, what if the responses of one group exhibit much more variance than the responses of another group? 

Let's first generate some data with heterogeneous variance: 

```{r}
# make example reproducible 
set.seed(1)

sample_size = 20
b0 = 1
b1 = 2
sd_residual = 0.5
mean_group1 = 1
sd_group1 = 1
mean_group2 = 30
sd_group2 = 10

df.variance = tibble(
  condition = rep(0:1, each = sample_size), 
  participant = rep(1:sample_size, 2)) %>% 
  group_by(participant) %>% 
  mutate(group = sample(1:2, size = 1),
         intercept = ifelse(group == 1,
                            rnorm(n(), mean = mean_group1, sd = sd_group1),
                            rnorm(n(), mean = mean_group2, sd = sd_group2))) %>% 
  group_by(condition) %>% 
  mutate(value = b0 + b1 * condition + intercept + rnorm(n(), sd = sd_residual)) %>% 
  ungroup %>% 
  mutate(condition = as.factor(condition),
         participant = as.factor(participant))
```

Let's fit the model: 

```{r}
# fit model
fit.variance = lmer(formula = value ~ 1 + group + condition + (1 | participant),
                data = df.variance)

fit.variance %>% summary()
```

Look at the data and model predictions: 

```{r}
fit.variance %>%
  augment() %>%
  clean_names() %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = value,
                       group = participant)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.5) +
  geom_point(aes(y = fitted),
             color = "red") +
  geom_line(aes(y = fitted),
             color = "red")
```

And the simulated data: 

```{r}
# simulated data 
fit.mixed %>%
  simulate() %>%
  bind_cols(df.mixed) %>%
  ggplot(data = .,
         mapping = aes(x = condition,
                       y = sim_1,
                       group = participant)) +
  geom_line(alpha = 0.5) +
  geom_point(alpha = 0.5)
```

The lmer() fails here. It uses one normal distribution to model the variance between participants. It cannot account for the fact that the answers of one groups of participants vary more than the answers from another groups of participants. Again, the simulated data doesn't look the original data, even though we did take the grouping into account. 

## Pooling and shrinkage 

Let's illustrate the concept of pooling and shrinkage via the sleep data set that comes with the lmer package. We've already loaded the data set into our environment as `df.sleep`. 

Let's start by visualizing the data 

```{r}
# visualize the data
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

The plot shows the effect of the number of days of sleep deprivation on the average reaction time (presumably in an experiment). Note that for participant 373 and 374 we only have one and two data points respectively. 

### Complete pooling 

Let's first fit a model the simply combines all the data points. This model ignores the dependence structure in the data (i.e. the fact that we have repeated observations from the same participants). 

```{r}
fit.complete = lm(formula = reaction ~ days,
                  data = df.sleep)

fit.params = tidy(fit.complete)

fit.complete %>% 
  summary()
```

And let's visualize the predictions of this model.

```{r}
# visualization (aggregate) 
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_abline(intercept = fit.params$estimate[1],
              slope = fit.params$estimate[2],
              color = "blue") +
  geom_point() +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

And here is what the model's predictions look like separated by participant.

```{r}
# visualization (separate participants) 
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_abline(intercept = fit.params$estimate[1],
              slope = fit.params$estimate[2],
              color = "blue") +
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

The model predicts the same relationship between sleep deprivation and reaction time for each participant (not surprising since we didn't even tell the model that this data is based on different participants). 

### No pooling 

We could also fit separate regressions for each participant. Let's do that.

```{r}
# fit regressions and extract parameter estimates 
df.no_pooling = df.sleep %>% 
  group_by(subject) %>% 
  nest(days, reaction) %>% 
  mutate(fit = map(data, ~ lm(reaction ~ days, data = .)),
         params = map(fit, tidy)) %>% 
  unnest(params) %>% 
  select(subject, term, estimate) %>% 
  complete(subject, term, fill = list(estimate = 0)) %>% 
  spread(term, estimate) %>% 
  clean_names()
```

And let's visualize what the predictions of these separate regressions would look like: 

```{r}
ggplot(data = df.sleep,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_abline(data = df.no_pooling %>% 
                filter(subject != 373),
              aes(intercept = intercept,
                  slope = days),
              color = "blue") +
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

When we fit separate regression, no information is shared between participants. 

### Partial pooling 

By usign linear mixed effects models, we are partially pooling information. That is, the estimates for one participant are influenced by the rest of the participants.

We'll fit a number of mixed effects models that differ in their random effects structure. 

#### Random intercept and random slope

This model allows for random differences in the intercepts and slopes between subjects (and also models the correlation between intercepts and slopes). 

Let's fit the model

```{r}
fit.random_intercept_slope = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                                  data = df.sleep)
```

and take a look at the model's predictions: 

```{r}
fit.random_intercept_slope %>% 
  augment() %>% 
  clean_names() %>% 
ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_line(aes(y = fitted),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

As we can see, the lines for each participant are different. We've allowed for the intercept as well as the relationship between sleep deprivation and reaction time to be different between participants. 

#### Only random intercepts 

Let's fit a model that only allows for the intercepts to vary between participants. 

```{r}
fit.random_intercept = lmer(formula = reaction ~ 1 + days + (1 | subject),
                            data = df.sleep)
```

And let's visualize what these predictions look like: 

```{r}
fit.random_intercept %>% 
  augment() %>% 
  clean_names() %>% 
ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_line(aes(y = fitted),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

Now, all the lines are parallel but the intercept differs between participants. 

#### Only random slopes 

Finally, let's compare a model that only allows for the slopes to differ but not the intercepts. 

```{r}
fit.random_slope = lmer(formula = reaction ~ 1 + days + (0 + days | subject),
                        data = df.sleep)
```

And let's visualize the model fit: 

```{r}
fit.random_slope %>% 
  augment() %>% 
  clean_names() %>% 
ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_line(aes(y = fitted),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

Here, all the lines have the same starting point (i.e. the same intercept) but the slopes are different. 

### Compare results 

Let's compare the results of the different methods -- complete pooling, no pooling, and partial pooling (with random intercepts and slopes). 

```{r, warning=F, message=F}
# complete pooling
fit.complete_pooling = lm(formula = reaction ~ days,
                          data = df.sleep)  

df.complete_pooling =  fit.complete_pooling %>% 
  augment() %>% 
  bind_rows(
    fit.complete_pooling %>% 
      augment(newdata = tibble(subject = c("373", "374"),
                               days = rep(10, 2)))
  ) %>% 
  clean_names() %>% 
  select(reaction, days, complete_pooling = fitted)

# no pooling
df.no_pooling = df.sleep %>% 
  group_by(subject) %>% 
  nest(days, reaction) %>% 
  mutate(fit = map(data, ~ lm(reaction ~ days, data = .)),
         augment = map(fit, augment)) %>% 
  unnest(augment) %>% 
  clean_names() %>% 
  select(subject, reaction, days, no_pooling = fitted)

# partial pooling
fit.lmer = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                data = df.sleep) 

df.partial_pooling = fit.lmer %>% 
  augment() %>% 
  bind_rows(
    fit.lmer %>% 
      augment(newdata = tibble(subject = c("373", "374"),
                               days = rep(10, 2)))
  ) %>% 
  clean_names() %>% 
  select(subject, reaction, days, partial_pooling = fitted)

# combine results
df.pooling = df.partial_pooling %>% 
  left_join(df.complete_pooling) %>% 
  left_join(df.no_pooling)
```

Let's compare the predictions of the different models visually: 

```{r}
ggplot(data = df.pooling,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_smooth(method = "lm",
              se = F,
              color = "orange",
              fullrange = T) + 
  geom_line(aes(y = complete_pooling),
            color = "green") + 
  geom_line(aes(y = partial_pooling),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

To better see the differences between the approaches, let's focus on the predictions for the participants with incomplete data: 

```{r}
# subselection
ggplot(data = df.pooling %>% 
         filter(subject %in% c("373", "374")),
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_smooth(method = "lm",
              se = F,
              color = "orange",
              fullrange = T) + 
  geom_line(aes(y = complete_pooling),
            color = "green") + 
  geom_line(aes(y = partial_pooling),
            color = "blue") + 
  geom_point() +
  facet_wrap(~subject) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

### Coefficients 

One good way to get a sense for what the different models are doing is by taking a look at the coefficients: 

```{r}
fit.complete_pooling %>% 
  coef()
```

```{r}
fit.random_intercept %>% 
  coef()
```

```{r}
fit.random_slope %>% 
  coef()
```

```{r}
fit.random_intercept_slope %>% 
  coef()
```

### Shrinkage 

In mixed effects models, the variance of parameter estimates across participants shrinks compared to a no pooling model (where we fit a different regression to each participant). Expressed differently, individual parameter estimates are borrowing strength from the overall data set in mixed effects models. 

```{r}
# get estimates from partial pooling model
df.partial_pooling = fit.random_intercept_slope %>% 
  coef() %>% 
  .[[1]] %>% 
  rownames_to_column("subject") %>% 
  clean_names()

# combine estimates from no pooling with partial pooling model 
df.plot = df.sleep %>% 
  group_by(subject) %>% 
  nest(days, reaction) %>% 
  mutate(fit = map(data, ~ lm(reaction ~ days, data = .)),
         tidy = map(fit, tidy)) %>% 
  unnest(tidy) %>% 
  select(subject, term, estimate) %>% 
  spread(term, estimate) %>% 
  clean_names() %>% 
  mutate(method = "no pooling") %>% 
  bind_rows(df.partial_pooling %>% 
              mutate(method = "partial pooling")) %>% 
  gather("index", "value", -c(subject, method)) %>% 
  mutate(index = factor(index, levels = c("intercept", "days")))

  
# visualize the results  
ggplot(data = df.plot,
       mapping = aes(x = value,
                     group = method,
                     fill = method)) + 
  stat_density(position = "identity",
               geom = "area",
               color = "black",
               alpha = 0.3) +
  facet_grid(cols = vars(index),
             scales = "free")
```

## Bootstrapping 

Bootstrapping is a good way to estimate our uncertainty on the parameter estimates in the model. 

### Linear model 

Let's briefly review how to do bootstrapping in a simple linear model. 

```{r}
# fit model 
fit.lm = lm(formula = reaction ~ 1 + days,
            data = df.sleep)

# coefficients
fit.lm %>% coef()

# bootstrapping 
df.boot = df.sleep %>% 
  bootstrap(n = 100,
            id = "id") %>% 
  mutate(fit = map(strap, ~ lm(formula = reaction ~ 1 + days, data = .)),
         tidy = map(fit, tidy)) %>% 
  unnest(tidy) %>% 
  select(id, term, estimate) %>% 
  spread(term, estimate) %>% 
  clean_names() 
```

Let's illustrate the linear model with a confidence interval (making parametric assumptions using the t-distribution). 

```{r}
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_smooth(method = "lm") + 
  geom_point(alpha = 0.3)
```

And let's compare this with the different regression lines that we get out of our bootstrapped samples:

```{r}
ggplot(data = df.sleep,
       mapping = aes(x = days, y = reaction)) + 
  geom_abline(data = df.boot,
              aes(intercept = intercept,
                  slope = days,
                  group = id),
              alpha = 0.1) +
  geom_point(alpha = 0.3)
```

#### bootmer() function

For the linear mixed effects model, we can use the `bootmer()` function to do bootstrapping. 

```{r, message=F, warning=F}
# fit the model 
fit.lmer = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                data = df.sleep)

# bootstrap parameter estimates 
boot.lmer = bootMer(fit.lmer,
                    FUN = fixef,
                    nsim = 100)

# compute confidence interval 
boot.ci(boot.lmer, index = 2, type = "perc")

# plot estimates 
boot.lmer$t %>% 
  as_tibble() %>% 
  clean_names() %>% 
  mutate(id = 1:n()) %>% 
  gather("index", "value", - id) %>% 
  ggplot(data = .,
       mapping = aes(x = value)) + 
  geom_density() + 
  facet_grid(cols = vars(index),
             scales = "free") +
  coord_cartesian(expand = F)
```

## Getting p-values 

We can use the "lmerTest" package to get p-values for the different fixed effects. 

```{r}
lmerTest::lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                data = df.sleep) %>% 
  summary()
```

## Understanding the lmer() syntax 

Here is an overview of how to specify different kinds of linear mixed effects models.

```{r, echo=F}
tibble(
  formula = c(
    "`dv ~ x1 + (1 | g)`",
    "`dv ~ x1 + (0 + x1 | g)`",
    "`dv ~ x1 + (x1 | g)`",
    "`dv ~ x1 + (x1 || g)`",
    "`dv ~ x1 + (1 | school) + (1 | teacher)`",
    "`dv ~ x1 + (1 | school/teacher)`"
  ),
  description = c(
    "Random intercept for each level of `g`",
    "Random slope for each level of `g`",
    "Correlated random slope and intercept for each level of `g`",
    "Uncorrelated random slope and intercept for each level of `g`",
    "Random intercept for each level of `school` and for each level of `teacher` (crossed)",
    "Random intercept for each level of `school` and for each level of `teacher` in `school` (nested)"
  )
) %>% 
  kable()
```

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r session}
sessionInfo()
```

## References