---
title: "Class 21"
author: "Tobias Gerstenberg"
date: ""
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=20"]
---

```{r setup, include=FALSE}
# these options here change the formatting of how comments are rendered
knitr::opts_chunk$set(comment = "#>",
                      fig.show = "hold")
```

# Bayesian data analysis

## Learning goals 

- Comparison between frequentist and Bayesian data analysis. 
- Objections to frequentist null hypothesis testing. 
- Benefits of Bayesian data analysis. 
- Bayesian models of cognition. 
- A simple worked coin flip example. 
- Posterior, prior, likelihood, sample size. 

## Load packages and set plotting theme  

```{r bda1-01, message=FALSE}
library("knitr")      # for knitting RMarkdown 
library("patchwork")  # for figure panels
library("tidyverse")  # for wrangling, plotting, etc. 
```

```{r bda1-02}
theme_set(theme_classic() + #set the theme 
    theme(text = element_text(size = 20))) #set the default text size
```

## Doing Bayesian inference "by hand"

### Sequential updating based on the Beta distribution 

```{r bda1-03}
# data 
data = c(0, 1, 1, 0, 1, 1, 1, 1)

# whether observation is a success or failure 
success = c(0, cumsum(data)) 
failure = c(0, cumsum(1 - data))
# I've added 0 at the beginning to show the prior

# plotting function
fun.plot_beta = function(success, failure){
  ggplot(data = tibble(x = c(0, 1)),
         mapping = aes(x = x)) +
    stat_function(fun = "dbeta",
                  args = list(shape1 = success + 1, shape2 = failure + 1),
                  geom = "area",
                  color = "black",
                  fill = "lightblue") +
    coord_cartesian(expand = F) +
    scale_x_continuous(breaks = seq(0.25, 0.75, 0.25)) + 
    theme(axis.title = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = margin(r = 1, t = 0.5, unit = "cm"))
}

# generate the plots 
plots = map2(success, failure, ~ fun.plot_beta(.x, .y))

# make a grid of plots
wrap_plots(plots, ncol = 3)
```

### Coin flip example 

Is the coin biased? 

```{r bda1-04}
# data 
data = rep(0:1, c(8, 2))

# parameters 
theta = c(0.1, 0.5, 0.9)

# prior 
prior = c(0.25, 0.5, 0.25)
# prior = c(0.1, 0.1, 0.8) # alternative setting of the prior
# prior = c(0.000001, 0.000001, 0.999998) # another prior setting 

# likelihood 
likelihood = dbinom(sum(data == 1), size = length(data), prob = theta)

# posterior 
posterior = likelihood * prior / sum(likelihood * prior)

# store in data frame 
df.coins = tibble(theta = theta,
                  prior = prior,
                  likelihood = likelihood,
                  posterior = posterior) 

```

Visualize the results: 

```{r bda1-05}
df.coins %>% 
  pivot_longer(cols = -theta,
               names_to = "index",
               values_to = "value") %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         theta = factor(theta, labels = c("p = 0.1", "p = 0.5", "p = 0.9"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       fill = index)) + 
  geom_bar(stat = "identity",
           color = "black") +
  facet_grid(rows = vars(index),
             switch = "y",
             scales = "free") + 
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line = element_blank())
```

### Bayesian inference by discretization

### Effect of the prior 

```{r bda1-06, fig.cap="Illustration of how the prior affects the posterior."}
# grid
theta = seq(0, 1, 0.01)

# data
data = rep(0:1, c(8, 2))

# calculate posterior
df.prior_effect = tibble(theta = theta, 
                  prior_uniform = dbeta(theta, shape1 = 1, shape2 = 1),
                  prior_normal = dbeta(theta, shape1 = 5, shape2 = 5),
                  prior_biased = dbeta(theta, shape1 = 8, shape2 = 2)) %>% 
  pivot_longer(cols = -theta,
               names_to = "prior_index",
               values_to = "prior") %>% 
  mutate(likelihood = dbinom(sum(data == 1),
                             size = length(data),
                             prob = theta)) %>% 
  group_by(prior_index) %>% 
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(theta, prior_index),
               names_to = "index",
               values_to = "value")

# make the plot
df.prior_effect %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         prior_index = factor(prior_index,
                              levels = c("prior_uniform", "prior_normal", "prior_biased"),
                              labels = c("uniform", "symmetric", "asymmetric"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       color = index)) +
  geom_line(size = 1) + 
  facet_grid(cols = vars(prior_index),
             rows = vars(index),
             scales = "free",
             switch = "y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank())
```

### Effect of the likelihood 

```{r bda1-07, fig.cap="Illustration of how the likelihood of the data affects the posterior."}
# grid
theta = seq(0, 1, 0.01)

df.likelihood_effect = tibble(theta = theta, 
                              prior = dbeta(theta, shape1 = 2, shape2 = 8),
                              likelihood_left = dbeta(theta, shape1 = 1, shape2 = 9),
                              likelihood_center = dbeta(theta, shape1 = 5, shape2 = 5),
                              likelihood_right = dbeta(theta, shape1 = 9, shape2 = 1)) %>% 
  pivot_longer(cols = -c(theta, prior),
               names_to = "likelihood_index",
               values_to = "likelihood") %>% 
  group_by(likelihood_index) %>% 
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(theta, likelihood_index),
               names_to = "index",
               values_to = "value")

df.likelihood_effect %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         likelihood_index = factor(likelihood_index,
                                   levels = c("likelihood_left", "likelihood_center", "likelihood_right"),
                                   labels = c("left", "center", "right"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       color = index)) +
  geom_line(size = 1) + 
  facet_grid(cols = vars(likelihood_index),
             rows = vars(index),
             scales = "free",
             switch = "y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank(),
        strip.text.x = element_blank())
  
```

### Effect of the sample size  

```{r bda1-08}
# grid
theta = seq(0, 1, 0.01)

df.sample_size_effect = tibble(theta = theta, 
                               prior = dbeta(theta, shape1 = 5, shape2 = 5),
                               likelihood_low = dbeta(theta, shape1 = 2, shape2 = 8),
                               likelihood_medium = dbeta(theta, shape1 = 10, shape2 = 40),
                               likelihood_high = dbeta(theta, shape1 = 20, shape2 = 80)) %>% 
  pivot_longer(cols = -c(theta, prior),
               names_to = "likelihood_index",
               values_to = "likelihood") %>% 
  group_by(likelihood_index) %>% 
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(theta, likelihood_index),
               names_to = "index",
               values_to = "value")

df.sample_size_effect %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         likelihood_index = factor(likelihood_index,
                                   levels = c("likelihood_low", "likelihood_medium", "likelihood_high"),
                                   labels = c("n = low", "n = medium", "n = high"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       color = index)) +
  geom_line(size = 1) + 
  facet_grid(cols = vars(likelihood_index),
             rows = vars(index),
             scales = "free",
             switch = "y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank())
```

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r bda1-09}
sessionInfo()
```
