---
title: "Class 21"
author: "Tobias Gerstenberg"
date: "March 5th, 2021"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=20"]
---

```{r, eval=FALSE, include=FALSE}
# Note: To get greta to work on my mac, I had to run the following commands 
devtools::install_github("greta-dev/greta")
greta::install_tensorflow(method = "conda",
                   version = "1.14.0",
                   extra_packages = "tensorflow-probability==0.7.0")
```

# Bayesian data analysis

## Learning goals 

- Doing Bayesian inference "by hand"
- Understanding the effect that prior, likelihood, and sample size have on the posterior. 
- Doing Bayesian data analysis with `greta`
  - A simple linear regression.

## Load packages and set plotting theme  

```{r, message=FALSE}
library("knitr")      # for knitting RMarkdown
library("janitor")    # for cleaning column names
library("patchwork")  # for figure panels
library("tidybayes")  # tidying up results from Bayesian models
library("greta")      # for writing Bayesian models
library("gganimate")  # for animations
library("extraDistr") # additional probability distributions
library("tidyverse")  # for wrangling, plotting, etc. 
```

```{r}
theme_set(theme_classic() + #set the theme 
    theme(text = element_text(size = 20))) #set the default text size

opts_chunk$set(comment = "",
               fig.show = "hold")
```

## Doing Bayesian inference "by hand"

### Sequential updating based on the Beta distribution 

```{r}
# data 
data = c(0, 1, 1, 0, 1, 1, 1, 1)

# whether observation is a success or failure 
success = c(0, cumsum(data)) 
failure = c(0, cumsum(1 - data))
# I've added 0 at the beginning to show the prior

# plotting function
fun.plot_beta = function(success, failure){
  ggplot(data = tibble(x = c(0, 1)),
         mapping = aes(x = x)) +
    stat_function(fun = "dbeta",
                  args = list(shape1 = success + 1, shape2 = failure + 1),
                  geom = "area",
                  color = "black",
                  fill = "lightblue") +
    coord_cartesian(expand = F) +
    scale_x_continuous(breaks = seq(0.25, 0.75, 0.25)) + 
    theme(axis.title = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          plot.margin = margin(r = 1, t = 0.5, unit = "cm"))
}

# generate the plots 
plots = map2(success, failure, ~ fun.plot_beta(.x, .y))

# make a grid of plots
wrap_plots(plots, ncol = 3)
```

### Coin flip example 

Is the coin biased? 

```{r}
# data 
data = rep(0:1, c(8, 2))

# parameters 
theta = c(0.1, 0.5, 0.9)

# prior 
prior = c(0.25, 0.5, 0.25)
# prior = c(0.1, 0.1, 0.8) # alternative setting of the prior
# prior = c(0.000001, 0.000001, 0.999998) # another prior setting 

# likelihood 
likelihood = dbinom(sum(data == 1), size = length(data), prob = theta)

# posterior 
posterior = likelihood * prior / sum(likelihood * prior)

# store in data frame 
df.coins = tibble(theta = theta,
                  prior = prior,
                  likelihood = likelihood,
                  posterior = posterior) 

```

Visualize the results: 

```{r}
df.coins %>% 
  pivot_longer(cols = -theta,
               names_to = "index",
               values_to = "value") %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         theta = factor(theta, labels = c("p = 0.1", "p = 0.5", "p = 0.9"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       fill = index)) + 
  geom_bar(stat = "identity",
           color = "black") +
  facet_grid(rows = vars(index),
             switch = "y",
             scales = "free") + 
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.title.x = element_blank(),
        axis.line = element_blank())
```

### Bayesian inference by discretization

#### Effect of the prior 

```{r, fig.cap="Illustration of how the prior affects the posterior."}
# grid
theta = seq(0, 1, 0.01)

# data
data = rep(0:1, c(8, 2))

# calculate posterior
df.prior_effect = tibble(theta = theta, 
                  prior_uniform = dbeta(theta, shape1 = 1, shape2 = 1),
                  prior_normal = dbeta(theta, shape1 = 5, shape2 = 5),
                  prior_biased = dbeta(theta, shape1 = 8, shape2 = 2)) %>% 
  pivot_longer(cols = -theta,
               names_to = "prior_index",
               values_to = "prior") %>% 
  mutate(likelihood = dbinom(sum(data == 1),
                             size = length(data),
                             prob = theta)) %>% 
  group_by(prior_index) %>% 
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(theta, prior_index),
               names_to = "index",
               values_to = "value")

# make the plot
df.prior_effect %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         prior_index = factor(prior_index,
                              levels = c("prior_uniform", "prior_normal", "prior_biased"),
                              labels = c("uniform", "symmetric", "asymmetric"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       color = index)) +
  geom_line(size = 1) + 
  facet_grid(cols = vars(prior_index),
             rows = vars(index),
             scales = "free",
             switch = "y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank())
```

#### Effect of the likelihood 

```{r, fig.cap="Illustration of how the likelihood of the data affects the posterior."}
# grid
theta = seq(0, 1, 0.01)

df.likelihood_effect = tibble(theta = theta, 
                              prior = dbeta(theta, shape1 = 2, shape2 = 8),
                              likelihood_left = dbeta(theta, shape1 = 1, shape2 = 9),
                              likelihood_center = dbeta(theta, shape1 = 5, shape2 = 5),
                              likelihood_right = dbeta(theta, shape1 = 9, shape2 = 1)) %>% 
  pivot_longer(cols = -c(theta, prior),
               names_to = "likelihood_index",
               values_to = "likelihood") %>% 
  group_by(likelihood_index) %>% 
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(theta, likelihood_index),
               names_to = "index",
               values_to = "value")

df.likelihood_effect %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         likelihood_index = factor(likelihood_index,
                                   levels = c("likelihood_left",
                                              "likelihood_center",
                                              "likelihood_right"),
                                   labels = c("left", "center", "right"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       color = index)) +
  geom_line(size = 1) + 
  facet_grid(cols = vars(likelihood_index),
             rows = vars(index),
             scales = "free",
             switch = "y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank(),
        strip.text.x = element_blank())
  
```

#### Effect of the sample size  

```{r}
# grid
theta = seq(0, 1, 0.01)

df.sample_size_effect = tibble(theta = theta, 
                               prior = dbeta(theta, shape1 = 5, shape2 = 5),
                               likelihood_low = dbeta(theta, shape1 = 2, shape2 = 8),
                               likelihood_medium = dbeta(theta,
                                                         shape1 = 10,
                                                         shape2 = 40),
                               likelihood_high = dbeta(theta,
                                                       shape1 = 20,
                                                       shape2 = 80)) %>% 
  pivot_longer(cols = -c(theta, prior),
               names_to = "likelihood_index",
               values_to = "likelihood") %>% 
  group_by(likelihood_index) %>% 
  mutate(posterior = likelihood * prior / sum(likelihood * prior)) %>% 
  ungroup() %>% 
  pivot_longer(cols = -c(theta, likelihood_index),
               names_to = "index",
               values_to = "value")

df.sample_size_effect %>% 
  mutate(index = factor(index, levels = c("prior", "likelihood", "posterior")),
         likelihood_index = factor(likelihood_index,
                                   levels = c("likelihood_low",
                                              "likelihood_medium",
                                              "likelihood_high"),
                                   labels = c("n = low", "n = medium", "n = high"))) %>% 
  ggplot(data = .,
         mapping = aes(x = theta,
                       y = value,
                       color = index)) +
  geom_line(size = 1) + 
  facet_grid(cols = vars(likelihood_index),
             rows = vars(index),
             scales = "free",
             switch = "y") +
  scale_x_continuous(breaks = seq(0, 1, 0.2)) +
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank())
```

## Doing Bayesian inference with Greta 

You can find out more about how get started with "greta" here: [https://greta-stats.org/articles/get_started.html](https://greta-stats.org/articles/get_started.html). Make sure to install the development version of "greta" (as shown in the "install-packages" code chunk above: `devtools::install_github("greta-dev/greta")`).

### Attitude data set 

```{r}
# load the attitude data set 
df.attitude = attitude
```

Visualize relationship between how well complaints are handled and the overall rating of an employee

```{r}
ggplot(data = df.attitude,
       mapping = aes(x = complaints,
                     y = rating)) +
  geom_point()
```

### Frequentist analysis 

```{r}
# fit model 
fit = lm(formula = rating ~ 1 + complaints, 
         data = df.attitude)

# print summary
fit %>% 
  summary()
```

Visualize the model's predictions

```{r}
ggplot(data = df.attitude,
       mapping = aes(x = complaints,
                     y = rating)) +
  geom_smooth(method = "lm",
              color = "black") + 
  geom_point()
```

### Bayesian regression

#### Fit the model

```{r}
set.seed(1)

# variables & priors
b0 = normal(0, 10)
b1 = normal(0, 10)
sd = cauchy(0, 3, truncation = c(0, Inf))

# linear predictor
mu = b0 + b1 * df.attitude$complaints

# observation model (likelihood)
distribution(df.attitude$rating) = normal(mu, sd)

# define the model
m = model(b0, b1, sd)
```

Visualize the model as graph: 

```{r}
# plotting
plot(m)
```

Draw samples from the posterior distribution: 

```{r, message=FALSE, cache=TRUE}
set.seed(1)

# sampling
draws = mcmc(m, n_samples = 1000)

# tidy up the draws
df.draws = tidy_draws(draws) %>% 
  clean_names()
```

#### Visualize the priors

These are the priors I used for the intercept, regression weights, and the standard deviation of the Gaussian likelihood function:  

```{r}
# Gaussian
ggplot(tibble(x = c(-30, 30)),
       aes(x = x)) +
  stat_function(fun = "dnorm", 
                size = 2,
                args = list(sd = 10))

# Cauchy
ggplot(tibble(x = c(0, 30)),
       aes(x = x)) +
  stat_function(fun = "dcauchy", 
                size = 2,
                args = list(location = 0,
                            scale = 3))
```

#### Visualize the posteriors

This is what the posterior looks like for the three parameters in the model: 

```{r}
df.draws %>% 
  select(draw:sd) %>% 
  pivot_longer(cols = -draw,
               names_to = "index",
               values_to = "value") %>% 
  ggplot(data = .,
         mapping = aes(x = value)) + 
  stat_density(geom = "line") + 
  facet_grid(rows = vars(index),
             scales = "free_y",
             switch = "y") + 
  annotate("segment", x = -Inf, xend = Inf, y = -Inf, yend = -Inf) + 
  annotate("segment", x = -Inf, xend = -Inf, y = -Inf, yend = Inf) + 
  theme(legend.position = "none",
        strip.background = element_blank(),
        axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 10),
        axis.line = element_blank(),
        strip.text.x = element_blank())
```

#### Visualize model predictions 

Let's take some samples from the posterior to visualize the model predictions: 

```{r}
ggplot(data = df.attitude,
       mapping = aes(x = complaints, 
                     y = rating)) + 
  geom_abline(data = df.draws %>% 
                sample_n(size = 50),
              mapping = aes(intercept = b0, 
                            slope = b1),
              alpha = 0.3,
              color = "lightblue") + 
  geom_point() 
```

#### Posterior predictive check 

Let's make an animation that illustrates what predicted data sets (based on samples from the posterior) would look like: 

```{r, message=FALSE}
p = df.draws %>% 
  sample_n(size = 10) %>%  
  mutate(complaints = list(seq(min(df.attitude$complaints),
                               max(df.attitude$complaints),
                               length.out = nrow(df.attitude)))) %>% 
  unnest(c(complaints)) %>% 
  mutate(prediction = b0 + b1 * complaints + rnorm(n(), sd = sd)) %>% 
  ggplot(aes(x = complaints, y = prediction)) + 
  geom_point(alpha = 0.8,
             color = "lightblue") +
  geom_point(data = df.attitude,
             aes(y = rating,
                 x = complaints)) +
  coord_cartesian(xlim = c(20, 100),
                  ylim = c(20, 100)) +
  transition_manual(draw)

animate(p,
        nframes = 60,
        width = 800,
        height = 600,
        res = 96,
        type = "cairo")

# anim_save("posterior_predictive.gif")
```

#### Prior predictive check 

And let's illustrate what data we would have expected to see just based on the information that we encoded in our priors. 

```{r, message=FALSE}
sample_size = 10

p = tibble(b0 = rnorm(sample_size, mean = 0, sd = 10),
           b1 = rnorm(sample_size, mean = 0, sd = 10),
           sd = rhcauchy(sample_size, sigma = 3),
           draw = 1:sample_size) %>% 
  mutate(complaints = list(runif(nrow(df.attitude),
                                 min = min(df.attitude$complaints),
                                 max = max(df.attitude$complaints)))) %>% 
  unnest(c(complaints)) %>% 
  mutate(prediction = b0 + b1 * complaints + rnorm(n(), sd = sd)) %>% 
  ggplot(aes(x = complaints, y = prediction)) + 
  geom_point(alpha = 0.8,
             color = "lightblue") +
  geom_point(data = df.attitude,
             aes(y = rating,
                 x = complaints)) +
  transition_manual(draw)

animate(p,
        nframes = 60,
        width = 800,
        height = 600,
        res = 96,
        type = "cairo")

# anim_save("prior_predictive.gif")
```

## Session info 

Information about this R session including which version of R was used, and what packages were loaded. 

```{r}
sessionInfo()
```
