---
title: "01_Introduction"
author: "Tobias Gerstenberg"
date: "1/7/2019"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

## Student feedback 

> I really liked the setup of going through the markdown file and doing examples together. It made very easy to jump ahead when we finish early and made it feel like a very efficient use of time

> I liked being able to work through the exercises. I wished we had more clear "learning goals" laid out at the start of the class.

> I think it may be more useful to have us to the practice exercises at home/outside of class and use class time to go over ideas/concepts etc...

> Writing all the packages we need to install in the Rmd file somewhere so we can just copy and paste into console, instead of installing along the way. I did appreciate it being in the email, but it's easier to have it in Rmd so that I don't have to have another window open

> Maybe telling everyone which line we are at now would help. Sometimes I lost track of where we were in the .Rmd file

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-2.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read_csv("how-was-the-pace-of-2.csv", skip = 12) %>%
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
ggsave("02_visualization1_pace.pdf", width = 8, height = 6)
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-2.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
ggsave("02_visualization1_overall.pdf", width = 8, height = 6)
```

## What to do next time

- explain the key components of an R Markdown file (have a slide on that before going into RStudio)
  - yaml header
  - headings 
  - text components 
  - code chunks 
  - inline code
  - how to execute code chunks
  - how to build a document 
- explain the basic setup of R Studio
- explain the + operator in `ggplot2`
- stress test materials 
  + the figures were missing and it took students a while to download the new materials
- make sure that my .Rmd file doesn't already have the plots pre-loaded 
- we only managed to get until section "Adding error bars" (incl. practice 2)





