---
title: "02_visualization1"
author: "Tobias Gerstenberg"
output: 
  html_document:
    theme: cosmo
    highlight: tango
---

```{r include=FALSE}
library("knitr")
library("DT")
library("janitor")
library("tidyverse")

opts_chunk$set(echo = TRUE)

theme_set(theme_classic() + 
            theme(text = element_text(size = 20),
                  panel.grid.major.y = element_line(color = "gray90")))
```

# Pace

```{r, echo=FALSE}
df.pace = read_csv("data/How was the pace of today's class-.csv",
                   skip = 1,
                   show_col_types = FALSE) %>% 
  clean_names() %>%  
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response,
        .drop = F)

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

# Happiness

```{r, echo=FALSE}
df.happy = read_csv("data/How happy were you with today's class overall-.csv",
                    skip = 1,
                    show_col_types = FALSE) %>% 
  clean_names() %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response,
        .drop = F)

ggplot(data = df.happy,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  scale_y_continuous(breaks = seq(0, 40, 2)) + 
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  theme(title = element_text(size = 18))
```

# What did you like? What can be improved? 

> stay on the new code longer before scrolling for those who might be typing out instead of pasting

> I liked diving in! Maybe slowing down a bit would be helpful for me, but I think that might be the case because I don't have a lot of R experience

> 	I think it would be more convenient if we can download class materials using Github fetch, which will keep our local class folder more organized. Thanks!

> 	apparently i had some package error so i was not able to participate. I thought i had downloaded everything prior to and even attended section, so i don't know if there was more I could have done to prepare, I didn't know there was an error

> I really liked the interactiveness from the breakout rooms working on a task together

```{r echo=FALSE, message=FALSE}
read_csv("data/What did you like about today's class- What could be improved next time-.csv",
         skip = 1,
         show_col_types = FALSE) %>% 
  datatable()
```

# What to do next time

- go slower in this first session in R 
- try not to jump around between different code sections as much 

