---
title: "20 Generalized linear model"
author: "Tobias Gerstenberg"
date: "2/26/2020"
output: 
  html_document:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
---

```{r setup, include=FALSE}
library("knitr")
library("DT")
library("tidyverse")
opts_chunk$set(echo = TRUE)
theme_set(
  theme_classic() + 
    theme(
      text = element_text(size = 20),
      panel.grid.major.y = element_line(color = "gray90")
    ) 
)
```

> Thank you for providing interpretations on the graphs as well as the issues that were problematic with them. It was a bit difficult to make jump into log odds without doing the math ourselves, but the slides lay it out well. I would still incorporate more time to talk to a neighbor for example when interpreting log odds.

> The classes on linear mixed effects and this class were great! I was able to understand things that I hadn't understood before. I particularly appreciated the material on how to interpret the coefficients of logistic regression! Thanks for these great classes, Tobi!

> I'd like to hear more explanations about how the concepts we learn are related to each other. For example, how is partial pooling related to linear mixed effects model?

> One thing in general that I feel like I still don't have a good sense of us how to model likert style data. I don't really have an idea of what a good strategy would be.

> I really appreciate learning about the mixed models and multilevel modeling, it is super helpful for research! __One thing that could be helpful in improving is giving a high-level definition of new concepts before jumping into the stats breakdown__. For example, "we use a logit transform to do X" before all of the complicated formulas appear, I got really lost there. (And also with concepts like AIC/BIC)

> Clear explanations on logistic regression. __Moving forward I really hope to know how we can address non-normality of residuals in our linear models.__

> I'm honestly a bit lost...

## Student feedback 

```{r echo=FALSE, message=FALSE}
read_csv("what-did-you-like-about-20.csv") %>% 
  set_names(c("response", "date")) %>% 
  select(response) %>% 
  datatable()
```

### PollEverywhere results 

```{r echo=FALSE, message=FALSE, results='hide'}
df.pace = read.csv("how-was-the-pace-of-20.csv", header = F) %>%
  filter(row_number() > which(.$V2 == "Created At")) %>% 
  set_names(c("response", "date")) %>% 
  mutate(response = factor(response, levels = c("much too slow", "a little too slow", "just right", "a little too fast", "much too fast"),
                           labels = c("much\ntoo slow", "a little\ntoo slow", "just right", "a little\ntoo fast", "much\ntoo fast"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.pace,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", fill = "lightblue", color = "black") +
  scale_y_continuous(breaks = seq(0, 30, 2), labels = seq(0, 30, 2)) + 
  labs(y = "number of students", x = "", title = "How was the pace of today's class?")
```

```{r echo=FALSE, message=FALSE, results='hide'}
df.overall = read_csv("how-happy-were-you-with-20.csv") %>% 
  set_names(c("response", "date")) %>% 
  filter(!str_detect(response, "Click")) %>% # remove response outside of the click regions 
  mutate(response = as.numeric(response),
         response = factor(response, levels = 1:5,
                           labels = c("very\nunhappy", "unhappy", "neutral", "happy", "very\nhappy"))) %>% 
  count(response) %>% 
  complete(response, fill = list(n = 0))

ggplot(data = df.overall,
       mapping = aes(x = response, y = n)) + 
  geom_bar(stat = "identity", aes(fill = response), color = "black", show.legend = F) +
  scale_fill_manual(values = c("red", "orange", "yellow", "lightgreen", "green")) +
  labs(y = "number of students", x = "", title = "How happy were you with today's class overall?") +
  scale_y_continuous(breaks = seq(0, 30, 2), labels = seq(0, 30, 2)) + 
  theme(title = element_text(size = 18))
```

## What to do next time

- try and have more interactivity in class again (e.g. talk to your neighbor exercises)
- go slower through the parts about the transformation to log odds 
- try to give high-level definitions of new concepts before then going into more detail in explaining them 
