---
title: "Class 23"
author: "Tobias Gerstenberg"
date: "March 6th, 2024"
output:
  bookdown::html_document2:
    toc: true
    toc_depth: 4
    theme: cosmo
    highlight: tango
    pandoc_args: ["--number-offset=22"]
---

# Bayesian data analysis 2

## Learning goals

- Building Bayesian models with `brms`.
	- Model evaluation: 
		- Visualizing and interpreting results. 
		- Testing hypotheses. 
	- Inference evaluation: Did things work out? 

## Load packages and set plotting theme

```{r, message=FALSE}
library("knitr")       # for knitting RMarkdown 
library("kableExtra")  # for making nice tables
library("janitor")     # for cleaning column names
library("tidybayes")   # tidying up results from Bayesian models
library("brms")        # Bayesian regression models with Stan
library("patchwork")   # for making figure panels
library("GGally")      # for pairs plot
library("broom.mixed") # for tidy lmer results
library("bayesplot")   # for visualization of Bayesian model fits 
library("modelr")      # for modeling functions
library("lme4")        # for linear mixed effects models 
library("afex")        # for ANOVAs
library("car")         # for ANOVAs
library("emmeans")     # for linear contrasts
library("ggeffects")   # for help with logistic regressions
library("titanic")     # titanic dataset
library("gganimate")   # for animations
library("parameters")  # for getting parameters
library("transformr")  # for gganimate
# install via: devtools::install_github("thomasp85/transformr")
library("tidyverse")   # for wrangling, plotting, etc. 
```

```{r}
theme_set(theme_classic() + # set the theme 
            theme(text = element_text(size = 20))) # set the default text size

opts_chunk$set(comment = "",
               fig.show = "hold")

options(dplyr.summarise.inform = F)

# set default color scheme in ggplot 
options(ggplot2.discrete.color = RColorBrewer::brewer.pal(9,"Set1"))
```

## Load data sets

```{r, message=FALSE}
# poker 
df.poker = read_csv("data/poker.csv") %>% 
  mutate(skill = factor(skill,
                        levels = 1:2,
                        labels = c("expert", "average")),
         skill = fct_relevel(skill, "average", "expert"),
         hand = factor(hand,
                       levels = 1:3,
                       labels = c("bad", "neutral", "good")),
         limit = factor(limit,
                        levels = 1:2,
                        labels = c("fixed", "none")),
         participant = 1:n()) %>% 
  select(participant, everything())

# sleep
df.sleep = sleepstudy %>% 
  as_tibble() %>% 
  clean_names() %>% 
  mutate(subject = as.character(subject)) %>% 
  select(subject, days, reaction) %>% 
  bind_rows(tibble(subject = "374",
                   days = 0:1,
                   reaction = c(286, 288)),
            tibble(subject = "373",
                   days = 0,
                   reaction = 245))

# titanic 
df.titanic = titanic_train %>% 
  clean_names() %>% 
  mutate(sex = as.factor(sex))

# politeness
df.politeness = read_csv("data/politeness_data.csv") %>% 
  rename(pitch = frequency)
```

## Poker

### 1. Visualize the data

Let's visualize the data first. 

```{r, warning=FALSE}
set.seed(1)

df.poker %>% 
  ggplot(mapping = aes(x = hand,
                       y = balance,
                       fill = hand,
                       group = skill,
                       shape = skill)) + 
  geom_point(alpha = 0.2,
             position = position_jitterdodge(dodge.width = 0.5,
                                             jitter.height = 0, 
                                             jitter.width = 0.2)) + 
  stat_summary(fun.data = "mean_cl_boot",
               position = position_dodge(width = 0.5),
               size = 1) + 
  labs(y = "final balance (in Euros)") + 
  scale_shape_manual(values = c(21, 22)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21,
                                                 fill = RColorBrewer::brewer.pal(3, "Set1"))),
         shape = guide_legend(override.aes = list(alpha = 1, fill = "black")))
```

### 2. Specify and fit the model

#### Frequentist model

And let's now fit a simple (frequentist) ANOVA model. You have multiple options to do so: 

```{r}
# Option 1: Using the "afex" package
aov_ez(id = "participant",
       dv = "balance",
       between = c("hand", "skill"),
       data = df.poker)

# Option 2: Using the car package (here we have to remember to set the contrasts to sum
# contrasts!)
lm(balance ~ hand * skill,
   contrasts = list(hand = "contr.sum",
                    skill = "contr.sum"),
   data = df.poker) %>% 
  car::Anova(type = 3)

# Option 3: Using the emmeans package (I like this one the best! It let's us use the 
# general lm() syntax and we don't have to remember to set the contrast)
fit.lm_poker = lm(balance ~ hand * skill,
                  data = df.poker) 

fit.lm_poker %>% 
  joint_tests()
```

All three options give the same result. Personally, I like Option 3 the best. 

#### Bayesian model

Now, let's fit a Bayesian regression model using the `brm()` function (starting with a simple model that only considers `hand` as a predictor):

```{r}
fit.brm_poker = brm(formula = balance ~ 1 + hand,
                    data = df.poker,
                    seed = 1, 
                    file = "cache/brm_poker")

# we'll use this model here later 
fit.brm_poker2 = brm(formula = balance ~ 1 + hand * skill,
                    data = df.poker,
                    seed = 1, 
                    file = "cache/brm_poker2")

fit.brm_poker %>%
  summary()
```

I use the `file = ` argument to save the model's results so that when I run this code chunk again, the model doesn't need to be fit again (fitting Bayesian models takes a while ...). And I used the `seed = ` argument to make this example reproducible. 

##### Full specification

So far, we have used the defaults that `brm()` comes with and not bothered about specifiying the priors, etc. 

Notice that we didn't specify any priors in the model. By default, "brms" assigns weakly informative priors to the parameters in the model. We can see what these are by running the following command: 

```{r}
fit.brm_poker %>% 
  prior_summary()
```

We can also get information about which priors need to be specified before fitting a model:

```{r}
get_prior(formula = balance ~ 1 + hand,
          family = "gaussian",
          data = df.poker)
```

Here is an example for what a more complete model specification could look like: 

```{r, message=FALSE}
fit.brm_poker_full = brm(formula = balance ~ 1 + hand,
                         family = "gaussian",
                         data = df.poker,
                         prior = c(prior(normal(0, 10),
                                         class = "b",
                                         coef = "handgood"),
                                   prior(normal(0, 10),
                                         class = "b",
                                         coef = "handneutral"),
                                   prior(student_t(3, 3, 10),
                                         class = "Intercept"),
                                   prior(student_t(3, 0, 10),
                                         class = "sigma")),
                         inits = list(list(Intercept = 0,
                                           sigma = 1,
                                           handgood = 5,
                                           handneutral = 5),
                                      list(Intercept = -5,
                                           sigma = 3,
                                           handgood = 2,
                                           handneutral = 2),
                                      list(Intercept = 2,
                                           sigma = 1,
                                           handgood = -1,
                                           handneutral = 1),
                                      list(Intercept = 1,
                                           sigma = 2,
                                           handgood = 2,
                                           handneutral = -2)),
                         iter = 4000,
                         warmup = 1000,
                         chains = 4,
                         file = "cache/brm_poker_full",
                         seed = 1)

fit.brm_poker_full %>%
  summary()
```

We can also take a look at the Stan code that the `brm()` function creates: 

```{r}
fit.brm_poker_full %>%
  stancode()
```

One thing worth noticing: by default, "brms" centers the predictors which makes it easier to assign a default prior over the intercept. 

### 3. Model evaluation

#### a) Did the inference work?

So far, we've assumed that the inference has worked out. We can check this by running `plot()` on our brm object:  

```{r, fig.height=8, fig.width=10}
plot(fit.brm_poker,
     N = 7,
     ask = F)
```

The posterior distributions (left hand side), and the trace plots of the samples from the posterior (right hand side) look good. 

Let's make our own version of a trace plot for one parameter in the model:

```{r}
fit.brm_poker %>% 
  spread_draws(b_Intercept) %>% 
  clean_names() %>% 
  mutate(chain = as.factor(chain)) %>% 
  ggplot(aes(x = iteration,
             y = b_intercept,
             group = chain,
             color = chain)) + 
  geom_line() + 
  scale_color_brewer(type = "seq",
                     direction = -1)
```

We can also take a look at the auto-correlation plot. Ideally, we want to generate independent samples from the posterior. So we don't want subsequent samples to be strongly correlated with each other. Let's take a look: 

```{r}
variables = fit.brm_poker %>%
  get_variables() %>%
  .[1:4]

fit.brm_poker %>% 
  as_draws() %>% 
  mcmc_acf(pars = variables,
           lags = 4)
```

Looking good! The autocorrelation should become very small as the lag increases (indicating that we are getting independent samples from the posterior). 

###### When things go wrong

Let's try to fit a model to very little data (just two observations) with extremely uninformative priors: 

```{r}
df.data = tibble(y = c(-1, 1))

fit.brm_wrong = brm(data = df.data,
                    family = gaussian,
                    formula = y ~ 1,
                    prior = c(prior(uniform(-1e10, 1e10), class = Intercept),
                              prior(uniform(0, 1e10), class = sigma)),
                    inits = list(list(Intercept = 0, sigma = 1),
                                 list(Intercept = 0, sigma = 1)),
                    iter = 4000,
                    warmup = 1000,
                    chains = 2,
                    file = "cache/brm_wrong")
```

Let's take a look at the posterior distributions of the model parameters: 

```{r}
summary(fit.brm_wrong)
```

Not looking good -- The estimates and credible intervals are off the charts. And the effective samples sizes in the chains are very small. 

Let's visualize the trace plots:

```{r, fig.height=6, fig.width=12}
plot(fit.brm_wrong,
     N = 2, 
     ask = F)
```

```{r}
fit.brm_wrong %>% 
  spread_draws(b_Intercept) %>% 
  clean_names() %>% 
  mutate(chain = as.factor(chain)) %>% 
  ggplot(aes(x = iteration,
             y = b_intercept,
             group = chain,
             color = chain)) + 
  geom_line() + 
  scale_color_brewer(direction = -1)
```

Given that we have so little data in this case, we need to help the model a little bit by providing some slighlty more specific priors. 

```{r}
fit.brm_right = brm(data = df.data,
                    family = gaussian,
                    formula = y ~ 1,
                    prior = c(prior(normal(0, 10), class = Intercept), # more reasonable priors
                              prior(cauchy(0, 1), class = sigma)),
                    iter = 4000,
                    warmup = 1000,
                    chains = 2,
                    seed = 1,
                    file = "cache/brm_right")
```

Let's take a look at the posterior distributions of the model parameters: 

```{r}
summary(fit.brm_right)
```

This looks much better. There is still quite a bit of uncertainty in our paremeter estimates, but it has reduced dramatically. 

Let's visualize the trace plots:

```{r}
plot(fit.brm_right,
     N = 2, 
     ask = F)
```

```{r}
fit.brm_right %>% 
  spread_draws(b_Intercept, sigma) %>% 
  clean_names() %>% 
  mutate(chain = as.factor(chain)) %>% 
  pivot_longer(cols = c(b_intercept, sigma)) %>% 
  ggplot(aes(x = iteration,
             y = value,
             group = chain,
             color = chain)) + 
  geom_line() + 
  facet_wrap(vars(name), ncol = 1) + 
  scale_color_brewer(direction = -1)
```

Looking mostly good!

#### b) Visualize model predictions

##### Posterior predictive check

To check whether the model did a good job capturing the data, we can simulate what future data the Bayesian model predicts, now that it has learned from the data we feed into it.  

```{r}
pp_check(fit.brm_poker, ndraws = 100)
```

This looks good! The predicted shaped of the data based on samples from the posterior distribution looks very similar to the shape of the actual data.  

Let's make a hypothetical outcome plot that shows what concrete data sets the model would predict.  The `add_predicted_draws()` function from the "tidybayes" package is helpful for generating predictions from the posterior.

```{r}
df.predictive_samples = df.poker %>% 
  add_predicted_draws(newdata = .,
                      object = fit.brm_poker2,
                      ndraws = 10)

p = ggplot(data = df.predictive_samples,
           mapping = aes(x = hand,
                         y = .prediction,
                         fill = hand,
                         group = skill,
                         shape = skill)) + 
  geom_point(alpha = 0.2,
             position = position_jitterdodge(dodge.width = 0.5,
                                             jitter.height = 0, 
                                             jitter.width = 0.2)) + 
  stat_summary(fun.data = "mean_cl_boot",
               position = position_dodge(width = 0.5),
               size = 1) + 
  labs(y = "final balance (in Euros)") + 
  scale_shape_manual(values = c(21, 22)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21)),
         shape = guide_legend(override.aes = list(alpha = 1, fill = "black"))) + 
  transition_manual(.draw)

animate(p, nframes = 120, width = 800, height = 600, res = 96, type = "cairo")
```

##### Prior predictive check

```{r}
fit.brm_poker_prior = brm(formula = balance ~ 0 + Intercept + hand * skill,
                          family = "gaussian",
                          data = df.poker,
                          prior = c(prior(normal(0, 10), class = "b"),
                                    prior(student_t(3, 0, 10), class = "sigma")),
                          iter = 4000,
                          warmup = 1000,
                          chains = 4,
                          file = "cache/brm_poker_prior",
                          sample_prior = "only",
                          seed = 1)

# generate prior samples 
df.prior_samples = df.poker %>% 
  add_predicted_draws(newdata = .,
                      object = fit.brm_poker_prior,
                      ndraws = 10)

# plot the results as an animation
p = ggplot(data = df.prior_samples,
           mapping = aes(x = hand,
                         y = .prediction,
                         fill = hand,
                         group = skill,
                         shape = skill)) + 
  geom_point(alpha = 0.2,
             position = position_jitterdodge(dodge.width = 0.5,
                                             jitter.height = 0, 
                                             jitter.width = 0.2)) + 
  stat_summary(fun.data = "mean_cl_boot",
               position = position_dodge(width = 0.5),
               size = 1) + 
  labs(y = "final balance (in Euros)") + 
  scale_shape_manual(values = c(21, 22)) + 
  guides(fill = guide_legend(override.aes = list(shape = 21,
                                                 fill = RColorBrewer::brewer.pal(3, "Set1"))),
         shape = guide_legend(override.aes = list(alpha = 1, fill = "black"))) + 
  transition_manual(.draw)

animate(p, nframes = 120, width = 800, height = 600, res = 96, type = "cairo")

# anim_save("poker_prior_predictive.gif")
```


### 4. Interpret the model parameters

#### Visualize the posteriors

Let's visualize what the posterior for the different parameters looks like. We use the `stat_halfeye()` function from the "tidybayes" package to do so: 


```{r, warning=FALSE}
fit.brm_poker %>% 
  as_draws_df() %>%
  select(starts_with("b_"), sigma) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(data = .,
         mapping = aes(y = fct_rev(variable),
                       x = value)) +
  stat_halfeye(fill = "lightblue") + 
  theme(axis.title.y = element_blank())
```

#### Compute highest density intervals

To compute the MAP (maximum a posteriori probability) estimate and highest density interval, we use the `mean_hdi()` function that comes with the "tidybayes" package.

```{r, warning=FALSE}
fit.brm_poker %>% 
  as_draws_df() %>%
  select(starts_with("b_"), sigma) %>% 
  mean_hdi() %>% 
  pivot_longer(cols = -c(.width:.interval),
               names_to = "index",
               values_to = "value") %>% 
  select(index, value) %>% 
  mutate(index = ifelse(str_detect(index, fixed(".")), index, str_c(index, ".mean"))) %>% 
  separate(index, into = c("parameter", "type"), sep = "\\.") %>% 
  pivot_wider(names_from = type, 
              values_from = value)
```

### 5. Test specific hypotheses

#### with `hypothesis()`

One key advantage of Bayesian over frequentist analysis is that we can test hypothesis in a very flexible manner by directly probing our posterior samples in different ways. 

We may ask, for example, what the probability is that the parameter for the difference between a bad hand and a neutral hand (`b_handneutral`) is greater than 0. Let's plot the posterior distribution together with the criterion: 

```{r,warning=FALSE}
fit.brm_poker %>% 
  as_draws_df() %>% 
  select(b_handneutral) %>% 
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(data = .,
         mapping = aes(y = variable, x = value)) +
  stat_halfeye(fill = "lightblue") + 
  geom_vline(xintercept = 0,
             color = "red")
```

We see that the posterior is definitely greater than 0. 

We can ask many different kinds of questions about the data by doing basic arithmetic on our posterior samples. The `hypothesis()` function makes this even easier. Here are some examples: 

```{r}
# the probability that the posterior for handneutral is less than 0
hypothesis(fit.brm_poker,
           hypothesis = "handneutral < 0")
```

```{r}
# the probability that the posterior for handneutral is greater than 4
hypothesis(fit.brm_poker,
           hypothesis = "handneutral > 4") %>% 
  plot()
```

```{r}
# the probability that good hands make twice as much as bad hands
hypothesis(fit.brm_poker,
           hypothesis = "Intercept + handgood > 2 * Intercept")
```

We can also make a plot of what the posterior distribution of the hypothesis looks like: 

```{r}
hypothesis(fit.brm_poker,
           hypothesis = "Intercept + handgood > 2 * Intercept") %>% 
  plot()
```


```{r}
# the probability that neutral hands make less than the average of bad and good hands
hypothesis(fit.brm_poker,
           hypothesis = "Intercept + handneutral < (Intercept + Intercept + handgood) / 2")
```

Let's double check one example, and calculate the result directly based on the posterior samples: 

```{r}
df.hypothesis = fit.brm_poker %>% 
  as_draws_df() %>% 
  clean_names() %>% 
  select(starts_with("b_")) %>% 
  mutate(neutral = b_intercept + b_handneutral,
         bad_good_average = (b_intercept + b_intercept + b_handgood)/2,
         hypothesis = neutral < bad_good_average)

df.hypothesis %>% 
  summarize(p = sum(hypothesis)/n())
```

#### with `emmeans()`

We can also use the `emmeans()` function to compute contrasts. 

```{r}
fit.brm_poker %>% 
  emmeans(specs = consec ~ hand)
```

Here, it computed the estimated means for each group for us, as well as the consecutive contrasts between each group. 

Let's visualize the contrasts. First, let's just use the `plot()` function as it's been adapted by the emmeans package: 

```{r}
fit.brm_poker %>% 
  emmeans(specs = consec ~ hand) %>% 
  pluck("contrasts") %>% 
  plot()
```

To get full posterior distributions instead of summaries, we can use the "tidybayes" package like so: 

```{r}
fit.brm_poker %>% 
  emmeans(specs = consec ~ hand) %>% 
  pluck("contrasts") %>% 
  gather_emmeans_draws() %>% 
  ggplot(mapping = aes(y = contrast,
                       x = .value)) + 
  stat_halfeye(fill = "lightblue",
               point_interval = mean_hdi,
               .width = c(0.5, 0.75, 0.95))
```



To see whether neutral hands did differently from bad and good hands (combined), we can define the following contrast.

```{r}
contrasts = list(neutral_vs_rest = c(-1, 2, -1))

fit.brm_poker %>% 
  emmeans(specs = "hand",
          contr = contrasts) %>% 
  pluck("contrasts") %>% 
  gather_emmeans_draws() %>% 
  mean_hdi()
```

Here, the HDP does not exclude 0. 

Let's double check that we get the same result using the `hypothesis()` function, or by directly computing from the posterior samples. 

```{r}
# using hypothesis()
fit.brm_poker %>% 
  hypothesis("(Intercept + handneutral)*2 < (Intercept + Intercept + handgood)")

# directly computing from the posterior
fit.brm_poker %>% 
  as_draws_df() %>% 
  clean_names() %>% 
  mutate(contrast = (b_intercept + b_handneutral) * 2 - (b_intercept + b_intercept + b_handgood)) %>% 
  summarize(contrast = mean(contrast))
```

The `emmeans()` function becomes particularly useful when our model has several categorical predictors, and we are interested in comparing differences along one predictor while marginalizing over the values of the other predictor. 

Let's take a look for a model that considers both `skill` and `hand` as predictors (as well as the interaction). 

```{r}
fit.brm_poker2 = brm(formula = balance ~ hand * skill,
                     data = df.poker,
                     seed = 1, 
                     file = "cache/brm_poker2")

fit.brm_poker2 %>% 
  summary()
```

In the summary table above, `skillexpert` captures the difference between an expert and an average player **when they have a bad hand**. To see whether there was a difference in expertise overall (i.e. across all three kinds of hands), we can calculate a linear contrast. 

```{r}
fit.brm_poker2 %>% 
  emmeans(pairwise ~ skill)
```

It looks like overall, skilled players weren't doing much better than average players. 

We can even do something like an equivalent of an ANOVA using `emmeans()`, like so: 

```{r}
joint_tests(fit.brm_poker2)
```

The values we get here are very similar to what we would get from a frequentist ANOVA: 

```{r}
aov_ez(id = "participant",
       dv = "balance",
       between = c("hand", "skill"),
       data = df.poker)
```

#### Bayes factor

Another way of testing hypothesis is via the Bayes factor. Let's fit the two models we are interested in comparing with each other: 

```{r, message=FALSE}
fit.brm_poker_bf1 = brm(formula = balance ~ 1 + hand,
                        data = df.poker,
                        save_pars = save_pars(all = T),
                        file = "cache/brm_poker_bf1")

fit.brm_poker_bf2 = brm(formula = balance ~ 1 + hand + skill,
                        data = df.poker,
                        save_pars = save_pars(all = T),
                        file = "cache/brm_poker_bf2")
```

And then compare the models using the `bayes_factor()` function: 

```{r}
bayes_factor(fit.brm_poker_bf2, fit.brm_poker_bf1)
```

Bayes factors don't have a very good reputation (see here and here). Instead, the way to go these days appears to be via approximate leave one out cross-validation. 

#### Approximate leave one out cross-validation

```{r}
fit.brm_poker_bf1 = add_criterion(fit.brm_poker_bf1,
                                  criterion = "loo",
                                  reloo = T,
                                  file = "cache/brm_poker_bf1")

fit.brm_poker_bf2 = add_criterion(fit.brm_poker_bf2,
                                  criterion = "loo",
                                  reloo = T,
                                  file = "cache/brm_poker_bf2")

loo_compare(fit.brm_poker_bf1,
            fit.brm_poker_bf2)
```


## Sleep study

### 1. Visualize the data

```{r}
set.seed(1)

ggplot(data = df.sleep %>% 
         mutate(days = as.factor(days)),
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(width = 0.1)) + 
  stat_summary(fun.data = "mean_cl_boot") 
```

### 2. Specify and fit the model

#### Frequentist analysis

```{r}
fit.lmer_sleep = lmer(formula = reaction ~ 1 + days + (1 + days | subject),
                      data = df.sleep)

fit.lmer_sleep %>% 
  summary()
```


#### Bayesian analysis

```{r}
fit.brm_sleep = brm(formula = reaction ~ 1 + days + (1 + days | subject),
                    data = df.sleep,
                    seed = 1,
                    file = "cache/brm_sleep")
```

### 3. Model evaluation

#### a) Did the inference work?

```{r, fig.height=16, fig.width=8}
fit.brm_sleep %>% 
  summary()

fit.brm_sleep %>% 
  plot(N = 6)
```

#### b) Visualize model predictions

```{r}
pp_check(fit.brm_sleep,
         ndraws = 100)
```

### 4. Interpret the parameters

```{r}
fit.brm_sleep %>% 
  tidy(conf.method = "HPDinterval")
```

#### Summary of posterior distributions

```{r, warning=FALSE}
# all parameters
fit.brm_sleep %>% 
  as_draws_df() %>% 
  select(-c(lp__, contains("["))) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(data = .,
         mapping = aes(x = value)) +
  stat_halfeye(point_interval = mode_hdi,
               fill = "lightblue") + 
  facet_wrap(~ variable,
             ncol = 2,
             scales = "free") +
  theme(text = element_text(size = 12))

# just the parameter of interest
fit.brm_sleep %>% 
  as_draws_df() %>% 
  select(b_days) %>%
  ggplot(data = .,
         mapping = aes(x = b_days)) +
  stat_halfeye(point_interval = mode_hdi,
               fill = "lightblue") + 
  theme(text = element_text(size = 12))
```

### 5. Test specific hypotheses

Here, we were just interested in how the number of days of sleep deprivation affected reaction time (and we can see that by inspecting the posterior for the `days` predictor in the model). 

### 6. Report results

#### Model prediction with posterior draws (aggregate)

```{r}
df.model = tibble(days = 0:9) %>% 
  add_linpred_draws(newdata = .,
                    object = fit.brm_sleep,
                    ndraws = 10,
                    seed = 1,
                    re_formula = NA)

ggplot(data = df.sleep,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(width = 0.1)) + 
  geom_line(data = df.model,
            mapping = aes(y = .linpred,
                          group = .draw),
            color = "lightblue") +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = 0:9)
```

#### Model prediction with credible intervals (aggregate)

```{r}
df.model = fit.brm_sleep %>% 
  fitted(re_formula = NA,
         newdata = tibble(days = 0:9)) %>% 
  as_tibble() %>% 
  mutate(days = 0:9) %>% 
  clean_names()
  
ggplot(data = df.sleep,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(width = 0.1)) + 
  geom_ribbon(data = df.model,
              mapping = aes(y = estimate,
                            ymin = q2_5,
                            ymax = q97_5),
              fill = "lightblue",
              alpha = 0.5) +
  geom_line(data = df.model,
            mapping = aes(y = estimate),
            color = "lightblue",
            size = 1) +
  stat_summary(fun.data = "mean_cl_boot") +
  scale_x_continuous(breaks = 0:9)
```


#### Model prediction with credible intervals (individual participants)

```{r, warning=FALSE, message=FALSE}
fit.brm_sleep %>% 
  fitted() %>% 
  as_tibble() %>% 
  clean_names() %>% 
  bind_cols(df.sleep) %>% 
  ggplot(data = .,
       mapping = aes(x = days,
                     y = reaction)) + 
  geom_ribbon(aes(ymin = q2_5,
                  ymax = q97_5),
              fill = "lightblue") +
  geom_line(aes(y = estimate),
            color = "blue") +
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

#### Model prediction for random samples

```{r}
df.model = df.sleep %>% 
  complete(subject, days) %>% 
  add_linpred_draws(newdata = .,
                    object = fit.brm_sleep,
                    ndraws = 10,
                    seed = 1)

df.sleep %>% 
  ggplot(data = .,
         mapping = aes(x = days,
                       y = reaction)) + 
  geom_line(data = df.model,
            aes(y = .linpred,
                group = .draw),
            color = "lightblue",
            alpha = 0.5) + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12))
```

#### Animated model prediction for random samples

```{r}
df.model = df.sleep %>% 
  complete(subject, days) %>% 
  add_linpred_draws(newdata = .,
                    object = fit.brm_sleep,
                    ndraws = 10,
                    seed = 1)

p = df.sleep %>% 
  ggplot(data = .,
         mapping = aes(x = days,
                       y = reaction)) + 
  geom_line(data = df.model,
            aes(y = .linpred,
                group = .draw),
            color = "black") + 
  geom_point() +
  facet_wrap(~subject, ncol = 5) +
  labs(x = "Days of sleep deprivation", 
       y = "Average reaction time (ms)") + 
  scale_x_continuous(breaks = 0:4 * 2) +
  theme(strip.text = element_text(size = 12),
        axis.text.y = element_text(size = 12)) + 
  transition_states(.draw, 0, 1) +
  shadow_mark(past = TRUE, alpha = 1/5, color = "gray50")

animate(p, nframes = 10, fps = 1, width = 800, height = 600, res = 96, type = "cairo")

# anim_save("sleep_posterior_predictive.gif")
```

## Titanic study

### 1. Visualize the data

```{r, message=FALSE}
df.titanic %>% 
  mutate(sex = as.factor(sex)) %>% 
  ggplot(data = .,
         mapping = aes(x = fare,
                       y = survived,
                       color = sex)) +
  geom_point(alpha = 0.1, size = 2) + 
  geom_smooth(method = "glm",
              method.args = list(family = "binomial"),
              alpha = 0.2,
              aes(fill = sex)) +
  scale_color_brewer(palette = "Set1")
```

### 2. Specify and fit the model

#### Frequentist analysis

```{r}
fit.glm_titanic = glm(formula = survived ~ 1 + fare * sex,
                      family = "binomial",
                      data = df.titanic)

fit.glm_titanic %>% 
  summary()
```

#### Bayesian analysis

```{r}
fit.brm_titanic = brm(formula = survived ~ 1 + fare * sex,
                      family = "bernoulli",
                      data = df.titanic,
                      file = "cache/brm_titanic",
                      seed = 1)
```

### 3. Model evaluation

#### a) Did the inference work?

```{r, fig.height=8, fig.width=10}
fit.brm_titanic %>% 
  summary()

fit.brm_titanic %>% 
  plot()
```

#### b) Visualize model predictions

```{r}
pp_check(fit.brm_titanic,
         ndraws = 100)
```

Let's visualize what the posterior predictive would have looked like for a linear model (instead of a logistic model). 

```{r}
fit.brm_titanic_linear = brm(formula = survived ~ 1 + fare * sex,
                             data = df.titanic,
                             file = "cache/brm_titanic_linear",
                             seed = 1)

pp_check(fit.brm_titanic_linear,
         ndraws = 100)
```

### 4. Interpret the parameters

```{r,warning=FALSE}
fit.brm_titanic %>% 
  as_draws_df() %>% 
  select(-lp__) %>%
  pivot_longer(cols = everything(),
               names_to = "variable",
               values_to = "value") %>% 
  ggplot(data = .,
         mapping = aes(y = variable,
                       x = value)) +
  stat_intervalh() + 
  scale_color_brewer()
```

```{r, eval=FALSE}
fit.brm_titanic %>% 
  parameters(centrality = "mean",
             ci = 0.95)
```

```{r, warning=F}
fit.brm_titanic %>% 
  ggpredict(terms = c("fare [0:500]", "sex")) %>% 
  plot()
```

### 5. Test specific hypotheses

Difference between men and women in survival? 

```{r}
fit.brm_titanic %>% 
  emmeans(specs = pairwise ~ sex,
          type = "response")
```

Difference in how fare affected the chances of survival for men and women? 

```{r}
fit.brm_titanic %>% 
  emtrends(specs = pairwise ~ sex,
           var = "fare")
```

### 6. Report results

```{r}
df.model = add_linpred_draws(newdata = expand_grid(sex = c("female", "male"),
                                                   fare = 0:500) %>% 
                               mutate(sex = factor(sex, levels = c("female", "male"))),
                             object = fit.brm_titanic,
                             ndraws = 10)
ggplot(data = df.titanic,
       mapping = aes(x = fare,
                     y = survived,
                     color = sex)) +
  geom_point(alpha = 0.1, size = 2) + 
  geom_line(data = df.model %>% 
              filter(sex == "male"),
            aes(y = .linpred,
                group = .draw,
                color = sex)) + 
  geom_line(data = df.model %>% 
              filter(sex == "female"),
            aes(y = .linpred,
                group = .draw,
                color = sex)) + 
  scale_color_brewer(palette = "Set1")
```

## Politeness data

The data is drawn from @winter2012phonetic, and this section follows the excellent tutorial by @franke2019bayesian.

(I'm skipping some of the steps of our recipe for Bayesian data analysis here.)

### 1. Visualize the data

```{r}
ggplot(data = df.politeness,
       mapping = aes(x = attitude,
                     y = pitch,
                     fill = gender,
                     color = gender)) + 
  geom_point(alpha = 0.2,
             position = position_jitter(width = 0.1, height = 0)) + 
  stat_summary(fun.data = "mean_cl_boot",
               shape = 21,
               size = 1,
               color = "black")
```

### 2. Specify and fit the model

#### Frequentist analysis

```{r}
fit.lm_polite = lm(formula = pitch ~ gender * attitude,
                   data = df.politeness)
```

#### Bayesian analysis

```{r}
fit.brm_polite = brm(formula = pitch ~ gender * attitude, 
                     data = df.politeness, 
                     file = "cache/brm_polite",
                     seed = 1)
```


### 5. Test specific hypotheses

#### Frequentist

```{r}
fit.lm_polite %>% 
  joint_tests()
```

It looks like there are significant main effects of gender and attitude, but no interaction effect. 

Let's check whether there is a difference in attitude separately for each gender: 

```{r}
fit.lm_polite %>% 
  emmeans(specs = pairwise ~ attitude | gender) %>% 
  pluck("contrasts")
```

There was a significant difference of attitude for female participants but not for male participants. 

#### Bayesian

Let's whether there was a main effect of gender.

```{r}
# main effect of gender
fit.brm_polite %>% 
  emmeans(specs = pairwise ~ gender) %>% 
  pluck("contrasts")

```

Let's take a look what the full posterior distribution over this contrast looks like: 

```{r}
fit.brm_polite %>% 
  emmeans(specs = pairwise ~ gender) %>% 
  pluck("contrasts") %>% 
  gather_emmeans_draws() %>% 
  ggplot(mapping = aes(x = .value)) + 
  stat_halfeye()
```

Looks neat! 

And let's confirm that we really estimated the main effect here. Let's fit a model that only has gender as a predictor, and then compare: 

```{r, warning=FALSE}
fit.brm_polite_gender = brm(formula = pitch ~ 1 + gender, 
                            data = df.politeness, 
                            file = "cache/brm_polite_gender",
                            seed = 1)

# using the gather_emmeans_draws to get means rather than medians 
fit.brm_polite %>% 
  emmeans(spec = pairwise ~ gender) %>% 
  pluck("contrasts") %>% 
  gather_emmeans_draws() %>% 
  mean_hdi()

fit.brm_polite_gender %>% 
  fixef() %>% 
  as_tibble(rownames = "term")
```

Yip, both of these methods give us the same result (the sign is flipped but that's just because emmeans computed F-M, whereas the other method computed M-F)! Again, the `emmeans()` route is more convenient because we can more easily check for several main effects (and take a look at specific contrast, too). 

```{r}
# main effect attitude
fit.brm_polite %>% 
  emmeans(specs = pairwise ~ attitude) %>% 
  pluck("contrasts")

# effect of attitude separately for each gender
fit.brm_polite %>% 
  emmeans(specs = pairwise ~ attitude | gender) %>% 
  pluck("contrasts")

# in case you want the means instead of medians 
fit.brm_polite %>% 
  emmeans(specs = pairwise ~ attitude | gender) %>% 
  pluck("contrasts") %>% 
  gather_emmeans_draws() %>% 
  mean_hdi()
```

Here is a way to visualize the contrasts: 

```{r}
fit.brm_polite %>% 
  emmeans(specs = pairwise ~ attitude | gender) %>% 
  pluck("contrasts") %>% 
  gather_emmeans_draws() %>% 
  ggplot(aes(x = .value,
             y = gender,
             fill = stat(x > 0))) + 
  facet_wrap(~ contrast) +
  stat_halfeye(show.legend = F) + 
  geom_vline(xintercept = 0, 
             linetype = 2) + 
  scale_fill_manual(values = c("gray80", "skyblue"))
```

Here is one way to check whether there was an interaction between attitude and gender (see [this vignette](https://cran.r-project.org/web/packages/emmeans/vignettes/interactions.html) for more info).

```{r}
fit.brm_polite %>% 
  emmeans(pairwise ~ attitude | gender) %>% 
  pluck("emmeans") %>% 
  contrast(interaction = c("consec"),
           by = NULL)
```


## Additional resources

- [Bayesian regression: Theory & Practice](https://michael-franke.github.io/Bayesian-Regression/)
- [Tutorial on visualizing brms posteriors with tidybayes](https://mjskay.github.io/tidybayes/articles/tidy-brms.html)
- [Hypothetical outcome plots](https://mucollective.northwestern.edu/files/2018-HOPsTrends-InfoVis.pdf)
- [Visual MCMC diagnostics](https://cran.r-project.org/web/packages/bayesplot/vignettes/visual-mcmc-diagnostics.html#general-mcmc-diagnostics)
- [Visualization of different MCMC algorithms](https://chi-feng.github.io/mcmc-demo/)
- [Article describing the different inference algorithms](https://elevanth.org/blog/2017/11/28/build-a-better-markov-chain/)

## Session info

Information about this R session including which version of R was used, and what packages were loaded.

```{r}
sessionInfo()
```
